---
title: "Group module GO enrichment + miranda + correlations"
author: "Jill Ashey"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(topGO)
library(igraph)
library(ggraph)
```

# Miranda and significant correlation data enrichment of miRNA targets

Read in modules from WGCNA combined with miranda data 
```{r}
all_info <- read.csv("../output/Molecular/miRNA_mRNA_interactions/gene_corr_miranda_PCC_sig_miRNA_mRNA.csv")
head(all_info)
```

Read in annotation file and make tab file for topGO
```{r}
annot <- read.delim("~/Desktop/GFFs/mcap/V3/Montipora_capitata_HIv3.genes.EggNog_results.txt")%>%
  dplyr::select(X.query, GOs) %>%
  mutate(across(everything(), ~ na_if(., "-")))

# Make sure all GO terms are separated by ;
annot$GOs <- gsub(",", ";", annot$GOs)

#write_tsv(annot, file = "../data/Molecular/Mcap_V3_gene2go.tab")
```

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Mcap_V3_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

# Maternal 

## Maternal - significant correlations

Perform GO enrichment analyses for genes with Maternal category Subset Maternal 
```{r}
maternal <- all_info %>%
  filter(category == "Maternal")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(maternal$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_BP_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_maternal$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_CC_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_maternal$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_MF_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_maternal$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_maternal, GO_CC_En_sig_gene_maternal, GO_MF_En_sig_gene_maternal)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_corr.pdf", last_plot(), width = 30, height = 40)
```

## Maternal - significant positive correlations

Perform GO enrichment analyses for Maternal genes with significant positive correlations with their target miRNAs.
```{r}
maternal <- all_info %>%
  filter(category == "Maternal") %>%
  filter(PCC.cor > 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(maternal$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_BP_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_maternal$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_CC_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_maternal$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_MF_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_maternal$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_maternal, GO_CC_En_sig_gene_maternal, GO_MF_En_sig_gene_maternal)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_positive_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_positive_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_positive_corr.pdf", last_plot(), width = 30, height = 40)
```

## Maternal - significant negative correlations

Perform GO enrichment analyses for Maternal genes with significant positive correlations with their target miRNAs.
```{r}
maternal <- all_info %>%
  filter(category == "Maternal") %>%
  filter(PCC.cor < 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(maternal$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_BP_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_maternal$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_CC_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_maternal$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with maternal df 
GO_MF_En_sig_gene_maternal <- maternal %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_maternal$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_maternal, GO_CC_En_sig_gene_maternal, GO_MF_En_sig_gene_maternal)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit() 

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_negative_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_negative_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_negative_corr.pdf", last_plot(), width = 30, height = 40)
```


# Minor Wave






# Major Wave 




















# MZT1 

## MZT1 - significant correlations

Perform GO enrichment analyses for genes with MZT1. category Subset MZT1 
```{r}
mzt1 <- all_info %>%
  filter(category == "MZT1")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_CC_En_sig_gene_mzt1<- mzt1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_MF_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt1, GO_CC_En_sig_gene_mzt1, GO_MF_En_sig_gene_mzt1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_corr.pdf", last_plot(), width = 30, height = 40)
```

## MZT1 - significant positive correlations

Perform GO enrichment analyses for genes MZT1 group that are positively correlated with miRNA targets. 
```{r}
mzt1 <- all_info %>%
  filter(category == "MZT1") %>%
  filter(PCC.cor > 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_CC_En_sig_gene_mzt1<- mzt1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_MF_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt1, GO_CC_En_sig_gene_mzt1, GO_MF_En_sig_gene_mzt1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_positive_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_positive_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_positive_corr.pdf", last_plot(), width = 30, height = 40)
```

## MZT1 - significant negative correlations

Perform GO enrichment analyses for genes MZT1 group that are negatively correlated with miRNA targets. 
```{r}
mzt1 <- all_info %>%
  filter(category == "MZT1") %>%
  filter(PCC.cor < 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_CC_En_sig_gene_mzt1<- mzt1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_MF_En_sig_gene_mzt1 <- mzt1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt1, GO_CC_En_sig_gene_mzt1, GO_MF_En_sig_gene_mzt1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_negative_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_negative_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt1_modules_miranda_sig_negative_corr.pdf", last_plot(), width = 30, height = 40)
```

# MZT2

## MZT2 - significant correlations

Perform GO enrichment analyses for genes with MZT2 category Subset MZT2
```{r}
mzt2 <- all_info %>%
  filter(category == "MZT2")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_CC_En_sig_gene_mzt2<- mzt2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt2, GO_CC_En_sig_gene_mzt2, GO_MF_En_sig_gene_mzt2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_corr.pdf", last_plot(), width = 30, height = 40)
```

## MZT2 - significant positive correlations

Perform GO enrichment analyses for genes in MZT2 category with positive correlations to their target miRNA
```{r}
mzt2 <- all_info %>%
  filter(category == "MZT2") %>%
  filter(PCC.cor > 0 )
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_CC_En_sig_gene_mzt2<- mzt2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt2, GO_CC_En_sig_gene_mzt2, GO_MF_En_sig_gene_mzt2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_positive_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_positive_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_positive_corr.pdf", last_plot(), width = 30, height = 40)
```

## MZT2 - significant negative correlations

Perform GO enrichment analyses for genes in MZT2 category with negative correlations to their target miRNA
```{r}
mzt2 <- all_info %>%
  filter(category == "MZT2") %>%
  filter(PCC.cor < 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(mzt2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt1 df 
GO_BP_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_mzt2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_CC_En_sig_gene_mzt2<- mzt2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_mzt2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_mzt2 <- mzt2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_mzt2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_mzt2, GO_CC_En_sig_gene_mzt2, GO_MF_En_sig_gene_mzt2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_negative_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_negative_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_mzt2_modules_miranda_sig_negative_corr.pdf", last_plot(), width = 30, height = 40)
```

# ZGA1

## ZGA1 - significant correlations

Perform GO enrichment analyses for genes with ZGA1 category Subset ZGA1
```{r}
zga1 <- all_info %>%
  filter(category == "ZGA1")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_BP_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_CC_En_sig_gene_zga1<- zga1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga1, GO_CC_En_sig_gene_zga1, GO_MF_En_sig_gene_zga1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_corr.pdf", last_plot(), width = 30, height = 40)
```

## ZGA1 - positive correlation

Perform GO enrichment analyses for genes in ZGA1 category that are positively correlated with their miRNA targets
```{r}
zga1 <- all_info %>%
  filter(category == "ZGA1") %>%
  filter(PCC.cor > 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_BP_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_CC_En_sig_gene_zga1<- zga1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga1, GO_CC_En_sig_gene_zga1, GO_MF_En_sig_gene_zga1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_positive_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_positive_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_positive_corr.pdf", last_plot(), width = 30, height = 40)
```

## ZGA1 - negative correlation

Perform GO enrichment analyses for genes in ZGA1 category that are negatively correlated with their miRNA targets
```{r}
zga1 <- all_info %>%
  filter(category == "ZGA1") %>%
  filter(PCC.cor < 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga1$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_BP_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga1$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga1 df 
GO_CC_En_sig_gene_zga1<- zga1 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga1$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
#annot <- annot %>%
#  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
#annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga1 <- zga1 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga1$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga1, GO_CC_En_sig_gene_zga1, GO_MF_En_sig_gene_zga1)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_negative_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_negative_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga1_modules_miranda_sig_negative_corr.pdf", last_plot(), width = 30, height = 40)
```

# ZGA2

## ZGA2 - significant correlations

Perform GO enrichment analyses for genes with ZGA2 category Subset ZGA2
```{r}
zga2 <- all_info %>%
  filter(category == "ZGA2")
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_BP_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_CC_En_sig_gene_zga2<- zga2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga2, GO_CC_En_sig_gene_zga2, GO_MF_En_sig_gene_zga2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_corr.pdf", last_plot(), width = 30, height = 40)
```

## ZGA2 - significant positive correlations

Perform GO enrichment analyses for genes in ZGA2 category that are positively associated with their target miRNAs
```{r}
zga2 <- all_info %>%
  filter(category == "ZGA2") %>%
  filter(PCC.cor > 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_BP_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_CC_En_sig_gene_zga2<- zga2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga2, GO_CC_En_sig_gene_zga2, GO_MF_En_sig_gene_zga2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_positive_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_positive_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_positive_corr.pdf", last_plot(), width = 30, height = 40)
```

## ZGA2 - significant negative correlations

Perform GO enrichment analyses for genes in ZGA2 category that are negatively associated with their target miRNAs
```{r}
zga2 <- all_info %>%
  filter(category == "ZGA2") %>%
  filter(PCC.cor < 0)
```

Make list of genes for input to topGO
```{r}
# Genes of interest 
clust_genes <- as.character(zga2$mRNA)

# All genes 
all_genes <- as.character(all_info$mRNA)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_BP_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_BP_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_BP_En_sig_gene_zga2$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with zga2 df 
GO_CC_En_sig_gene_zga2<- zga2 %>%
  inner_join(GO_CC_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_CC_En_sig_gene_zga2$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Join with mzt2 df 
GO_MF_En_sig_gene_zga2 <- zga2 %>%
  inner_join(GO_MF_En_sig_gene, by = c("mRNA" = "X.query"))

# Add ontology column 
GO_MF_En_sig_gene_zga2$ontology <- "Molecular Functions"
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene <- rbind(GO_BP_En_sig_gene_zga2, GO_CC_En_sig_gene_zga2, GO_MF_En_sig_gene_zga2)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene <- GO_En_sig_gene %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
write.csv(GO_En_sig_gene, "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_negative_corr.csv")
```

Plot 
```{r}
plot_data <- GO_En_sig_gene %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA),
            Fisher = mean(Fisher), .groups = 'drop') %>%
  group_by(ontology) %>%
  slice_max(order_by = gene_count, n = 10)

ggplot(plot_data, aes(x = reorder(Term, gene_count), y = gene_count, fill = Fisher)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = gene_count), vjust = -0.5, color = "black", size = 3) +
  coord_flip() +
  scale_fill_gradient(low = "blue", high = "red") +
  facet_grid(vars(ontology), scales = "free", space = "free_y") + 
  labs(x = "GO Term",
       y = "Number of Genes",
       fill = "Fisher Value") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 36, face = "bold"),        # Axis title size
    axis.text = element_text(size = 34, colour = "black"),                        # Axis text size
    legend.title = element_text(size = 34, face = "bold"),      # Legend title size
    legend.text = element_text(size = 32),                      # Legend text size
    strip.text = element_text(size = 34, face = "bold"),        # Facet text size
    strip.background = element_rect(fill = "lightgray", color = "black", size = 1.5),         # Facet background
    axis.line = element_line(size = 1, colour = "black"),       # Enhanced axis lines
    axis.ticks = element_line(size = 1),                        # Thicker axis ticks
     #panel.border = element_blank()                             # Remove panel border
    #panel.grid.major = element_blank(),                         # Remove major grid lines
    #panel.grid.minor = element_blank()                          # Remove minor grid lines
        panel.border = element_rect(color = "black", size = 1.2),   # Enhanced facet border lines
    panel.grid.major = element_line(size = 0.5, color = "gray"), # Grid lines inside the facets
    panel.spacing = unit(1, "lines"),                           # Increase space between facets
    strip.placement = "outside"    
  )

ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_negative_corr.png", last_plot(), width = 30, height = 40)
ggsave(filename = "../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_zga2_modules_miranda_sig_negative_corr.pdf", last_plot(), width = 30, height = 40)
```


# Assess GO terms

Read in annot info
```{r}
annot <- read.delim("~/Desktop/GFFs/mcap/V3/Montipora_capitata_HIv3.genes.EggNog_results.txt")
```

Investigate enriched functions of maternal significant positive correlations 
```{r}
maternal_sig_pos_corr <- read.csv("../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_positive_corr.csv")
length(unique(maternal_sig_pos_corr$miRNA))
length(unique(maternal_sig_pos_corr$mRNA))
length(unique(maternal_sig_pos_corr$Term))

blah <- maternal_sig_pos_corr %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA))

maternal_sig_pos_corr_annot <- maternal_sig_pos_corr %>%
  inner_join(annot, by = c("mRNA" = "X.query"))
length(unique(maternal_sig_pos_corr_annot$Preferred_name))
unique(maternal_sig_pos_corr_annot$Preferred_name)
```

Investigate enriched functions of maternal significant negative correlations 
```{r}
maternal_sig_neg_corr <- read.csv("../output/Molecular/miRNA_mRNA_interactions/enrichment_miranda_corr/GO_en_sig_maternal_modules_miranda_sig_negative_corr.csv")
length(unique(maternal_sig_neg_corr$miRNA))
length(unique(maternal_sig_neg_corr$mRNA))
length(unique(maternal_sig_neg_corr$Term))

blah <- maternal_sig_neg_corr %>%
  filter(ontology != "Cellular Components") %>%
  group_by(Term, ontology) %>%
  summarise(gene_count = n_distinct(mRNA))

maternal_sig_neg_corr_annot <- maternal_sig_neg_corr %>%
  inner_join(annot, by = c("mRNA" = "X.query"))
length(unique(maternal_sig_neg_corr_annot$Preferred_name))
unique(maternal_sig_neg_corr_annot$Preferred_name)
```

Plot the preferred names of maternal negative correlations
```{r}
plot_data <- maternal_sig_neg_corr_annot %>%
  distinct(Preferred_name, p_value, PCC.cor) %>%
  arrange(p_value)

# Create the plot
ggplot(plot_data, aes(x = p_value, y = reorder(Preferred_name, -p_value), color = PCC.cor)) +
  geom_point(size = 3) +
  #scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  #scale_x_log10() +  # Use log scale for p-values
  theme_bw() +
  labs(x = "p-value", 
       y = "", 
       color = "PCC correlation")+
  theme(axis.text.y = element_text(size = 8))  # Adjust text size if needed

# Save the plot if desired
# ggsave("gene_correlations_plot.png", width = 10, height = 12, dpi = 300)

```




