---
title: "lncRNA - mRNA interactions"
author: "Jill Ashey"
date: "2024-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


Looking at lncRNA mRNA correlations 
```{r}
lcount <- read.csv("../output/Molecular/lncRNA/lncRNA_filtered_counts.csv")[1:10,]
rownames(lcount) <- lcount[,1] #set first column that contains gene names as rownames
lcount <- lcount[,-1] # remove column w/ gene names 
lcount_1hpf <- lcount[,c("M6", "M7", "M8", "M9")]

mcount <- read.csv("../output/Molecular/mRNA/mRNA_filtered_counts.csv")[1:10,]
rownames(mcount) <- mcount[,1] #set first column that contains gene names as rownames
mcount <- mcount[,-1] # remove column w/ gene names 
mcount_1hpf <- mcount[,c("M6", "M7", "M8", "M9")]

lcount_numeric <- apply(lcount_1hpf, 2, as.numeric)
mcount_numeric <- apply(mcount_1hpf, 2, as.numeric)

calculate_pcc <- function(lncRNA, mRNA) {
  result <- cor.test(lncRNA, mRNA, method = "pearson")
  return(c(pcc = result$estimate, p_value = result$p.value))
}

results <- list()
for (i in 1:nrow(lcount_numeric)) {
  lncRNA <- lcount_numeric[i,]
  for (j in 1:nrow(mcount_numeric)) {
    mRNA <- mcount_numeric[j,]
    results[[paste(rownames(lcount)[i], rownames(mcount)[j], sep = "_")]] <- calculate_pcc(lncRNA, mRNA)
  }
}

results_df <- do.call(rbind, results)
results_df <- as.data.frame(results_df)
results_df$lncRNA_mRNA <- rownames(results_df)
rownames(results_df) <- NULL

#co_expressed_pairs <- results_df[abs(results_df$pcc) > 0.5 & results_df$p_value < 0.05, ]
co_expressed_pairs <- results_df[results_df$p_value < 0.05, ]
```

Huang et al. 2017 used abs value of correlation as 0.95
Lopes et al. 2019 used 0.8
Yin et al. 2020 (oral cancer) used top 0.05 percentile

NEED TO SEPARATE BY LIFE STAGE
also maybe need to vst or rlog counts data to normalize 

A positive correlation between an mRNA and an lncRNA indicates a strong relationship between their expression levels. This means that as the expression of one increases, the expression of the other tends to increase as well. Here's what this correlation suggests:
Co-expression: The mRNA and lncRNA are likely to be co-expressed, meaning they are transcribed together or in response to similar stimuli12.
Regulatory Relationship: The lncRNA may be involved in regulating the expression of the mRNA, or vice versa. This could occur through various mechanisms:
Transcriptional regulation: The lncRNA might influence the transcription of the mRNA gene3.
Post-transcriptional regulation: The lncRNA could affect mRNA stability or processing3.
Epigenetic regulation: The lncRNA might be involved in modifying the chromatin state around the mRNA gene2.
Functional Association: The correlated expression suggests that the lncRNA and mRNA may be involved in related biological processes or pathways4.

Plot a correlation example 
```{r}
lncRNA_name <- "Montipora_capitata_HIv3___Scaffold_7:10436852-10446721"
lncRNA_expression <- lcount_1hpf[lncRNA_name, , drop = FALSE]

mRNA_name <- "Montipora_capitata_HIv3___RNAseq.g8204.t1"
mRNA_expression <- mcount_1hpf[mRNA_name, , drop = FALSE]

test <- rbind(lncRNA_expression, mRNA_expression)

test <- tibble::rownames_to_column(test, var = "ID")

test_long <- pivot_longer(
  test,
  cols = starts_with("M"),   # Select all columns starting with "M"
  names_to = "Sample",       # New column for the variable names
  values_to = "Value"        # New column for the values
)

comparison_data <- test_long %>%
  pivot_wider(names_from = ID, values_from = Value)

colnames(comparison_data) <- c("Sample", "lncRNA", "mRNA")

ggplot(comparison_data, aes(x = lncRNA, y = mRNA, color = Sample)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    x = "Montipora_capitata_HIv3___Scaffold_190:50628-63004",
    y = "Montipora_capitata_HIv3___TS.g35745.t1",
    title = "Comparison of Values Between Two IDs"
  ) +
  theme_minimal()
```




