---
title: "Temperature Data"
author: "jillashey"
date: "2023-07-03"
output: html_document
---

This script reads and plots temperature data from Hobo loggers (tidbits and U22s) for the developmental timeseries project in Hawaii 2023.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Load packages
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
library(emmeans)
library(data.table)

getwd()
```

## Read in Tidbit Hobo logger files 
```{r}
tidbit.files <- list.files(path="../data/environmental/logger_mcap", pattern="*27.csv", full.names=TRUE)

tidbits <- rbindlist(lapply(tidbit.files, function(file) {
  dt <- fread(file, select = c("Date-Time (HST)", "Ch: 1 - Temperature   (°C)"))
  dt[, logger := sub(".*/", "", file)]
  dt[, logger := sub("_.*", "", logger)]
  setnames(dt, c("DateTime", "temp.C", "logger"))
  dt[!is.na(temp.C)]
}))

tidbits$DateTime<-as.POSIXct(tidbits$DateTime, format="%m/%d/%Y %H:%M:%S", tz="Pacific/Honolulu")
```

## Read in U22 Hobo logger files 
```{r}
u22.files <- list.files(path="../data/environmental/logger_mcap", pattern="*30.csv", full.names=TRUE)

u22 <- rbindlist(lapply(u22.files, function(file) {
  # Only read necessary columns to save memory
  dt <- fread(file, select = c("Date Time, GMT-10:00", "Temp, °C"))
  # Add logger ID and clean names
  dt[, logger := sub("_.*", "", sub(".*/", "", file))]
  setnames(dt, c("DateTime", "temp.C", "logger"))
  # Remove NAs and return
  dt[!is.na(temp.C)]
})) %>% as_tibble()

u22$DateTime<-as.POSIXct(u22$DateTime, format="%m/%d/%y %H:%M", tz="Pacific/Honolulu")
```

## Bind u22 and tidbit loggers
```{r}
loggers <- rbind(tidbits, u22)

# check logger SNs
levels(as.factor(loggers$logger)) ## 8 loggers, as expected. 2 per incubator and 4 per treatment
```

## Truncate dates so that date range is only capturing the experimental period
```{r}
loggers <- loggers%>%
  filter(DateTime < as.POSIXct("2023-06-20 21:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-17 22:30:00"))
```

## Read in treatment information that assigns the loggers to a specific treatment
```{r}
metadata <- read.csv("../data/environmental/logger_meta_mcap.csv")
metadata$logger <- as.character(metadata$logger)
```

## Assign treatment information to logger
```{r}
loggers <- full_join(loggers, metadata, by = "logger")
loggers$Type <- as.factor(loggers$Type)
loggers$Incubator <- as.factor(loggers$Incubator)
loggers$Treatment <- as.factor(loggers$Treatment)
#loggers$DateTime <- as.numeric(loggers$DateTime)
```

## Plot temperature by incubator colored by treatment 
```{r}
temp_plot1<-loggers%>%
  ggplot(aes(x=DateTime, y=temp.C, colour=Treatment))+
  geom_point()+
  #ylim(26, 30)+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  theme_classic(); temp_plot1
```

## Summarize by treatment with error shading for standard error
```{r}
temp_plot_treatment1<-loggers%>%
  group_by(Treatment, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sem, lower=mean-sem)%>%
  
  ggplot(aes(x=DateTime, y=mean, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE) +
  geom_vline(xintercept=as.POSIXct("2023-06-17 23:30:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 02:45:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 09:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 14:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 21:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-19 02:30:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-19 21:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-20 21:00:00"), linetype="dashed", color="black")+
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  #xlab("Date Time")+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red"), name="Temperature")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14)
    #legend.position="none"
  ); temp_plot_treatment1

# Save plot 
ggsave("../output/environmental/logger_temperature_smooth.png", temp_plot_treatment1, width=6, height=4)
```

## Calculate the mean difference in temperature treatments between groups.  
```{r}
loggers%>%
  group_by(Treatment, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE))%>%
  spread(key=Treatment, value=mean)%>%
  reframe(difference=High-Ambient)%>%
  reframe(mean=mean(difference))
# mean difference between high and ambient is 3.784998
```

## Calculate mean temperature for each incubator.      
```{r}
loggers %>%
  mutate(day = floor_date(DateTime, "day")) %>%
           group_by(Incubator, Treatment) %>%
           summarise(mean = mean(temp.C))
```

## Subset and plot ambient treatment only 
```{r}
amb_loggers <- loggers %>%
  filter(Treatment == "Ambient")

temp_plot_amb<-amb_loggers%>%
  group_by(DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sem, lower=mean-sem)%>%
  ggplot(aes(x=DateTime, y=mean))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, color = "blue"), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour="blue"), alpha=1, show.legend=TRUE) +
  geom_vline(xintercept=as.POSIXct("2023-06-17 23:30:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 02:45:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 09:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 14:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 21:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-19 02:30:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-19 21:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-20 21:00:00"), linetype="dashed", color="black")+
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  xlab("")+
  #xlab("Date Time")+
  scale_colour_manual(values=c("blue"), name="Temperature")+
  scale_fill_manual(values=c("blue"), name="Temperature")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14),
    legend.position="none"
  ); temp_plot_amb

# Save plot 
ggsave("../output/environmental/ambient_logger_temperature_smooth.png", temp_plot_amb, width=6, height=4) 
ggsave("../output/environmental/ambient_logger_temperature_smooth.pdf", temp_plot_amb, width=6, height=4)
```

Calculate mean and SEM for each ambient incubator
```{r}
amb_means<-amb_loggers%>%
  group_by(Incubator)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sem, lower=mean-sem)
```



## Make plot for ICRS proposal 
In my 2024 ICRS application, I am proposing to examine 4 timepoints: 9, 22, 48, and 72 hpf. I need to adjust my temperature plot to only show those timepoints.

Summarize by treatment with error shading for standard error
```{r}
temp_plot_treatment_icrs<-loggers%>%
  group_by(Treatment, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sd, lower=mean-sd)%>%
  
  ggplot(aes(x=DateTime, y=mean, colour=Treatment, fill=Treatment))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=Treatment), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=Treatment), alpha=1, show.legend=TRUE) +
  geom_vline(xintercept=as.POSIXct("2023-06-17 23:30:00"), linetype="solid", color="black")+ 
  #geom_vline(xintercept=as.POSIXct("2023-06-18 02:45:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 09:00:00"), linetype="dashed", color="black")+ 
  #geom_vline(xintercept=as.POSIXct("2023-06-18 14:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-18 21:00:00"), linetype="dashed", color="black")+ 
  #geom_vline(xintercept=as.POSIXct("2023-06-19 02:30:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-19 21:00:00"), linetype="dashed", color="black")+ 
  geom_vline(xintercept=as.POSIXct("2023-06-20 21:00:00"), linetype="dashed", color="black")+
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  xlab("Date")+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red"), name="Temperature")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14)
    #legend.position="none"
  ); temp_plot_treatment_icrs

# Save plot 
ggsave("../figures/environmental/logger_temperature_smooth_icrs.png", temp_plot_treatment_icrs, width=6, height=4)
```





