---
title: "miRNA discovery - mirdeep2"
author: "Jill Ashey"
date: "2024-11-18"
output: html_document
---

I ran mirdeep2 on R1 reads from DT Mcap 2023 experiment. The output was files with novel and known miRNAs. In this code, I filter the results to identify putative miRNAs!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
```

## Semi raw reads 
NOTE: these miRNAs are from semi-raw reads - I still need to trim adapters from certain reads. 

Load novel miRNAs 
```{r}
novel <- read.csv("../data/Molecular/smRNA/mirdeep2/all_novel.csv", header = T)

length(unique(novel$provisional.id)) # 752 unique genomic ids
length(unique(novel$consensus.mature.sequence)) # 625 unique miRNA sequences
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. 
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(novel_filt$provisional.id))
length(unique(novel_filt$consensus.mature.sequence))

# Save csv 
#write.csv(novel_filt, "../data/Molecular/smRNA/mirdeep2/novel_filt.csv")
```

Load the known miRNAs
```{r}
known <- read.csv("../data/Molecular/smRNA/mirdeep2/all_known.csv")
length(unique(known$tag.id)) # 14 unique genomic ids
length(unique(known$consensus.mature.sequence)) # 14 unique miRNA sequences
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(known_filt$tag.id))
length(unique(known_filt$consensus.mature.sequence))

# Save csv 
#write.csv(known_filt, "../data/Molecular/smRNA/mirdeep2/known_filt.csv")
```

## Reads trimmed to 75bp with flexbar 

Load novel miRNAs 
```{r}
novel_75 <- read.csv("../output/Molecular/smRNA/mirdeep2/flexbar_75bp/flexbar_75bp_novel.csv", header = T)

length(unique(novel_75$provisional.id)) # 610 unique genomic ids
length(unique(novel_75$consensus.mature.sequence)) # 507 unique miRNA sequences
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. 
```{r}
novel_75_filt <- novel_75 %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(novel_75_filt$provisional.id))
length(unique(novel_75_filt$consensus.mature.sequence))

# Save csv 
#write.csv(novel_75_filt, "../output/Molecular/smRNA/mirdeep2/flexbar_75bp/flexbar_75bp_novel_filt.csv")
```

Load the known miRNAs
```{r}
known_75 <- read.csv("../output/Molecular/smRNA/mirdeep2/flexbar_75bp/flexbar_75bp_known.csv", header = T)
length(unique(known_75$tag.id)) # 11 unique genomic ids
length(unique(known_75$consensus.mature.sequence)) # 11 unique miRNA sequences
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_75_filt <- known_75 %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(known_75_filt$tag.id))
length(unique(known_75_filt$consensus.mature.sequence))

# Save csv 
#write.csv(known_75_filt, "../output/Molecular/smRNA/mirdeep2/flexbar_75bp/flexbar_75bp_known_filt.csv")
```

## Reads trimmed to 35bp with flexbar 

Load novel miRNAs 
```{r}
novel_35 <- read.csv("../output/Molecular/smRNA/mirdeep2/flexbar_35bp/flexbar_35bp_novel.csv", header = T)

length(unique(novel_35$provisional.id)) # 592 unique genomic ids
length(unique(novel_35$consensus.mature.sequence)) # 497 unique miRNA sequences
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. 
```{r}
novel_35_filt <- novel_35 %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(novel_35_filt$provisional.id))
length(unique(novel_35_filt$consensus.mature.sequence))

# Save csv 
write.csv(novel_35_filt, "../output/Molecular/smRNA/mirdeep2/flexbar_35bp/flexbar_35bp_novel_filt.csv")
```

Load the known miRNAs
```{r}
known_35 <- read.csv("../output/Molecular/smRNA/mirdeep2/flexbar_35bp/flexbar_35bp_known.csv", header = T)
length(unique(known_35$tag.id)) # 10 unique genomic ids
length(unique(known_35$consensus.mature.sequence)) # 10 unique miRNA sequences
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_35_filt <- known_35 %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(known_35_filt$tag.id))
length(unique(known_35_filt$consensus.mature.sequence))

# Save csv 
write.csv(known_35_filt, "../output/Molecular/smRNA/mirdeep2/flexbar_35bp/flexbar_35bp_known_filt.csv")
```

## Trim stringent (cutadapt) first batch novel and known miRNAs 

Load novel miRNAs 
```{r}
novel <- read.csv("../output/Molecular/smRNA/mirdeep2/trim_stringent_first_batch/trim_stringent_first_batch_novel.csv", header = T)

length(unique(novel$provisional.id)) # 659 unique genomic ids
length(unique(novel$consensus.mature.sequence)) # 554 unique miRNA sequences
```

Filter so that mirdeep2 score >10, no rfam info, at least 10 reads in mature and star read count, and significant randfold p-value. 
```{r}
novel_filt <- novel %>%
  filter(miRDeep2.score > 10) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(novel_filt$provisional.id))
length(unique(novel_filt$consensus.mature.sequence))

# Save csv 
write.csv(novel_filt, "../output/Molecular/smRNA/mirdeep2/trim_stringent_first_batch/trim_stringent_first_batch_novel_filt.csv")
```

Load the known miRNAs
```{r}
known <- read.csv("../output/Molecular/smRNA/mirdeep2/trim_stringent_first_batch/trim_stringent_first_batch_known.csv", header = T)
length(unique(known$tag.id)) # 15 unique genomic ids
length(unique(known$consensus.mature.sequence)) # 15 unique miRNA sequences
```

Because we can see some mapping to known miRNAs in cnidarians, I'm going to filter so that mirdeep2 score >0, no rfam info, at least 10 reads in mature read count, and significant ranfold p-value
```{r}
known_filt <- known %>%
  filter(miRDeep2.score > 0) %>%
  filter(rfam.alert == "-") %>%
  filter(mature.read.count > 10) %>%
  #filter(star.read.count >10) %>%
  filter(significant.randfold.p.value == "yes") #%>%
  #na.omit()
length(unique(known_filt$tag.id))
length(unique(known_filt$consensus.mature.sequence))

# Save csv 
write.csv(known_filt, "../output/Molecular/smRNA/mirdeep2/trim_stringent_first_batch/trim_stringent_first_batch_known_filt.csv")
```

