---
title: "mirtarrnaseq"
author: "Jill Ashey"
date: "2024-12-27"
output: html_document
---

Running mirtarrnaseq - following part 2 of their [vignette](https://www.bioconductor.org/packages/release/bioc/vignettes/mirTarRnaSeq/inst/doc/mirTarRnaSeq.pdf).
.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(mirTarRnaSeq)
library(reshape2)
library(igraph)
library(ggraph)
library(tidygraph)

#options(stringsAsFactors = FALSE) #Set Strings to character
```

The format for the miRNA and mRNA files should be: gene/miRNA as column 1, log2FC as column 2, and pvalue as column 3. I am comparing all timepoints to 1hpf (the first timepoint). 

Read in mRNA file and format. I have 7 files to read in that represent each timepoint compared to 1hpf. 
```{r}
file_path <- "../output/Molecular/mRNA/"

# Get a list of all files matching the pattern
file_list <- list.files(path = file_path, pattern = "^mRNA_DE_results_all_1v", full.names = TRUE)

# Read in file, select and rename columns, store in list 
data_list <- lapply(file_list, function(file) {
  df <- read.csv(file)
  df_selected <- df[, c("X", "log2FoldChange", "padj")]
  names(df_selected) <- c("gene", "log2FoldChange", "pvalue")
  return(df_selected)
})

names(data_list) <- tools::file_path_sans_ext(basename(file_list))
```

Get mRNAs with particular fold change
```{r}
mrna_data <- one2OneRnaMiRNA(data_list, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)

mrna_data_fc <- one2OneRnaMiRNA(data_list, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)$foldchanges

mrna_data_p <- one2OneRnaMiRNA(data_list, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)$pvalues
```
The `one2OneRnaMiRNA` function grabs the gene name, FC and pvalue information from each df to produce a set of genes that are significant in any of the gene lists. It then filters each df to extract the FC and pvalues of the significant genes present in all input dfs. 

Read in miRNA file and format. 
```{r}
file_path_mi <- "../output/Molecular/smRNA/"

# Get a list of all files matching the pattern
file_list_mi <- list.files(path = file_path_mi, pattern = "^miRNA_DE_results_all_1v", full.names = TRUE)

# Read in file, select and rename columns, store in list 
data_list_mi <- lapply(file_list_mi, function(file) {
  df <- read.csv(file)
  df_selected <- df[, c("X", "log2FoldChange", "padj")]
  names(df_selected) <- c("gene", "log2FoldChange", "pvalue")
  return(df_selected)
})

names(data_list_mi) <- tools::file_path_sans_ext(basename(file_list_mi))
```

Get mRNAs with particular fold change
```{r}
mirna_data <- one2OneRnaMiRNA(data_list_mi, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)

mirna_data_fc <- one2OneRnaMiRNA(data_list_mi, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)$foldchanges

mirna_data_p <- one2OneRnaMiRNA(data_list_mi, gene_colname = "gene", fc_colname = "log2FoldChange", pthreshold = 0.05)$pvalues
```

Get mRNA miRNA correlation 
```{r}
# subset mrna data 
mrna_subset <- mrna_data_fc[1:20, , drop = FALSE]
rownames(mrna_subset) <- rownames(mrna_data_fc)[1:20]

corr0 <- corMirnaRna(mrna_subset, mirna_data_fc, method = "pearson")

# all data
#corr0 <- corMirnaRna(mrna_data_fc, mirna_data_fc, method = "pearson")
```
This function is calculating the Pearson correlation coefficient for each miRNA mRNA combination. Pearson correlation coefficient (the value in the resulting df) is calculated by taking the sum of the products of the differences of each data point from their respective means, divided by the product of the standard deviations of each variable.

I need to subset the data because the memory exceeds what my computer can do when running this function. 

Subset the mRNA data as chunks and run the chunks instead of the whole df
```{r}
# Function to process a chunk
process_chunk <- function(mrna_chunk, mirna_data) {
  corr_chunk <- corMirnaRna(mrna_chunk, mirna_data, method = "pearson")
  return(corr_chunk)
}

# Set chunk size
chunk_size <- 1000

# Calculate number of chunks
num_chunks <- ceiling(nrow(mrna_data_fc) / chunk_size)

# Initialize list to store results
corr_list <- list()

# Store original miRNA rownames
mirna_rownames <- rownames(mirna_data_fc)

# Process data in chunks
for (i in 1:num_chunks) {
  start_idx <- (i-1) * chunk_size + 1
  end_idx <- min(i * chunk_size, nrow(mrna_data_fc))
  
  mrna_chunk <- mrna_data_fc[start_idx:end_idx, , drop = FALSE]
  
  # Ensure row names are set for both dataframes
  rownames(mrna_chunk) <- rownames(mrna_data_fc)[start_idx:end_idx]
  
  corr_chunk <- process_chunk(mrna_chunk, mirna_data_fc)
  
  corr_list[[i]] <- corr_chunk
  
  cat("Processed chunk", i, "of", num_chunks, "\n")
}

# Combine results
corr_0 <- do.call(rbind, corr_list)

# Replace miRNA_1, miRNA_2, etc. with original miRNA names
corr_0$V1 <- rep(mirna_rownames, nrow(mrna_data_fc))

# Display the first few rows of the result
print(head(corr_0))

```



Make background distribution correlation
```{r}
#outs <- sampCorRnaMirna(mrna_data_fc, mirna_data_fc, Shrounds = 100, Srounds = 1000)
```

Similarly to above, I have so much mRNA data that R is struggling to handle it with memory. I need to break into chunks again 
```{r}
chunk_size <- 1000  # Adjust based on your system's memory capacity
num_chunks <- ceiling(nrow(mrna_data_fc) / chunk_size)

# Function to process a chunk
process_chunk <- function(mrna_chunk, mirna_data, Shrounds, Srounds) {
  sampCorRnaMirna(mrna_chunk, mirna_data, Shrounds = Shrounds, Srounds = Srounds)
}

# Initialize list to store results
outs_list <- list()

# Process data in chunks
for (i in 1:num_chunks) {
  start_idx <- (i-1) * chunk_size + 1
  end_idx <- min(i * chunk_size, nrow(mrna_data_fc))
  
  mrna_chunk <- mrna_data_fc[start_idx:end_idx, , drop = FALSE]
  
  outs_chunk <- process_chunk(mrna_chunk, mirna_data_fc, Shrounds = 100, Srounds = 1000)
  
  outs_list[[i]] <- outs_chunk
  
  cat("Processed chunk", i, "of", num_chunks, "\n")
}

# Combine results
outs <- do.call(rbind, outs_list)
```

Make density plots for background and correlation in our data 
```{r}
mirRnaDensityCor(corr0, outs)
```

Grey is background distribution and red is our data. 

```{r}
sig_corrs <- threshSig(corr_0, outs, pvalue = 0.05)
```

Read in miranda data
```{r}
miranda <- read.delim("../output/Molecular/smRNA/mirdeep2/miranda_strict_all_1kb_parsed_mcap.txt", header = F)
colnames(miranda) <- c("miRNA", "mRNA", "score", "energy", "query_start_end", "subject_start_end", "total_bp_shared", "query_similar", "subject_similar")

# Format miranda df 
miranda$miRNA <- sub("^>", "", miranda$miRNA)  # Remove leading ">"
miranda$mRNA <- sub("ID=", "", miranda$mRNA)
miranda$mRNA <- sub("::.*", "", miranda$mRNA)
```

Select specific columns and rename for the package to work
```{r}
miranda <- miranda %>%
  dplyr::select(miRNA, mRNA, score, energy, query_similar, subject_similar) %>%
  dplyr::rename(V1 = miRNA, V2 = mRNA, V3 = score, V4 = energy, V5 = query_similar, V6 = subject_similar)
```

Correlation heatmap 
```{r}
# Function to process a chunk
process_chunk <- function(mrna_chunk, mirna_data, threshold, miranda) {
  corMirnaRnaMiranda(mrna_chunk, mirna_data, threshold, miranda)
}

# Set chunk size
chunk_size <- 1000

# Calculate number of chunks
num_chunks <- ceiling(nrow(mrna_data_fc) / chunk_size)

# Initialize list to store results
corr_list <- list()

# Process data in chunks
for (i in 1:num_chunks) {
  start_idx <- (i-1) * chunk_size + 1
  end_idx <- min(i * chunk_size, nrow(mrna_data_fc))
  
  mrna_chunk <- mrna_data_fc[start_idx:end_idx, , drop = FALSE]
  
  corr_chunk <- process_chunk(mrna_chunk, mirna_data_fc, -0.7, miranda)
  
  corr_list[[i]] <- corr_chunk
  
  cat("Processed chunk", i, "of", num_chunks, "\n")
}

# Combine results
newcorr <- do.call(rbind, corr_list)

# Plot and save
pdf("../output/Molecular/miRNA_mRNA_interactions/mirtarrnaseq_correlations.pdf", height = 20, width = 25)
mirRnaHeatmap(newcorr)
dev.off()

# Save data as csv
write.csv(newcorr, "../output/Molecular/miRNA_mRNA_interactions/mirtarrnaseq_correlations_data.csv")
```

Plot using igraph
```{r}
subset <- newcorr[1:100,]

# Step 1: Create the graph from the data frame
g <- graph_from_data_frame(subset[, c("V1", "V2", "value")], directed = FALSE)

# Step 2: Add edge attributes
E(g)$weight <- abs(E(g)$value)  # Use absolute PCC for edge weight
E(g)$color <- ifelse(E(g)$value > 0, "blue", "red")  # Blue for positive, red for negative correlations

# Step 3: Add node attributes
V(g)$type <- ifelse(V(g)$name %in% subset$V1, "miRNA", "mRNA")

# Convert to tbl_graph for ggraph
g_tbl <- as_tbl_graph(g)

# Step 4: Create the plot
ggraph(g_tbl, layout = "fr") +
  geom_edge_link(aes(edge_width = weight, color = color), alpha = 0.6) +
  geom_node_point(aes(color = type), size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width(range = c(0.5, 3)) +
  scale_color_manual(values = c("miRNA" = "lightblue", "mRNA" = "lightgreen", 
                                 "Positive correlation" = "blue", 
                                 "Negative correlation" = "red")) +
  theme_graph() +
  labs(title = "miRNA-mRNA Interaction Network",
       subtitle = "Edge width represents |PCC|, color represents correlation direction")

```



Read in mRNA count data
```{r}
mrna_counts <- read.csv("../output/Molecular/mRNA/mRNA_filtered_counts.csv")

# Select specific columns 
mrna_counts <- mrna_counts %>%
  select(X, M9, M13, M23, M35, M52, M60, M72, M85)

# Rename cols based on hpf 
colnames(mrna_counts) <- c("gene", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf", "28_hpf", "48_hpf", "72_hpf")
```

Read in miRNA count data 
```{r}
mirna_counts <- read.delim("../output/Molecular/smRNA/mirdeep2/miRNAs_expressed_all_samples_1733514422.csv", header = T)
mirna_counts <- as.data.frame(mirna_counts)

# Select specific columns 
mirna_counts <- mirna_counts %>%
  select(X.miRNA, s09.norm., s13.norm., s23.norm., s35.norm., s52.norm., s60.norm., s72.norm., s85.norm.)

# Rename cols based on hpf 
colnames(mirna_counts) <- c("mirna", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf", "28_hpf", "48_hpf", "72_hpf")
```

Join count with correlation data and save 
```{r}
new_df <- newcorr %>%
  full_join(mirna_counts, by = c("V1" = "mirna"))

new_df <- new_df %>%
  full_join(mrna_counts, by = c("V2" = "gene"))

write.csv(new_df, "../output/Molecular/miRNA_mRNA_interactions/mirtarrnaseq_correlations_counts_data.csv")
```

Subset data and plot 
```{r}
newcorr_one <- newcorr %>%
  filter(V1 == c("Montipora_capitata_HIv3___Scaffold_14_1018693"))

mirna_subset <- mirna_counts[mirna_counts$mirna %in% newcorr_one$V1, ]
mrna_subset <- mrna_counts[mrna_counts$gene %in% newcorr_one$V2, ]
```

```{r}
mirna_melted <- melt(mirna_subset, id.vars = "mirna", variable.name = "hpf", value.name = "count")
mirna_melted <- mirna_melted %>% dplyr::rename(gene = mirna) 
mirna_melted$type <- "miRNA"  # Add a column to indicate type

# Melt the mRNA data
mrna_melted <- melt(mrna_subset, id.vars = "gene", variable.name = "hpf", value.name = "count")
mrna_melted$type <- "mRNA"  # Add a column to indicate type

# Combine both melted dataframes
combined_data <- rbind(mirna_melted, mrna_melted)

# Merge combined_data with interaction information
ggplot(combined_data, aes(x = hpf, y = count, group = gene)) +
  geom_line(aes(color = gene)) + 
  geom_point(aes(color = gene)) +
  scale_x_discrete(limits = c("1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf", "28_hpf", "48_hpf", "72_hpf")) +
  scale_y_log10() +
  labs(x = "Hours Post Fertilization (hpf)", y = "Count (log scale)", 
       title = "miRNA and mRNA Interactions Over Time") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14),
    legend.position = "right",
    legend.text = element_text(size = 8)
  )
```

Need to make better plots









