---
title: "mRNA polyA differential expression"
author: "Jill Ashey"
date: "2025-04-29"
output: html_document
---

This script provides details on mRNA polyA differential expression across time point in Mcap development.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# The following setting is important, do not omit.
options(stringsAsFactors = FALSE) #Set Strings to character
```

Load packages 
```{r}
library(genefilter)
library(DESeq2)
library(pheatmap)
library(lme4)
library(tidyverse)
library(car)
library(gplots)
library(wesanderson)
library(cowplot)
library(gridExtra)
library(grid)
library("factoextra")
library(vegan)
library(ggcorrplot)
```

Load counts data
```{r}
counts <- read.csv("../data/Molecular/mRNA_polyA/Mcap_polyA_gene_count_matrix.csv")
counts <- as.data.frame(counts)
rownames(counts) <- counts[,1] #set first column that contains gene names as rownames
counts <- counts[,-1] # remove column w/ gene names 
```

We have 32 columns as expected. Remove extra info from the column names 
```{r}
colnames(counts) <- gsub(".bam.gtf", "", colnames(counts))
```

Remove any genes that have 0 counts across all samples (ie these genes were not expressed)
```{r}
dim(counts) 

counts<-counts %>%
     mutate(Total = rowSums(.[, 1:32]))%>%
    filter(!Total==0)%>%
    dplyr::select(!Total)

dim(counts)
```

Started with 54384 genes, the total number in the annotation. After removing genes with 0s across all samples, 27366 genes are left. 

Filter data with pOverA. There are 32 samples with an n=4 samples per timepoint. Genes will pass filtering if they are present in 4/32=0.125 of the samples because we expect differential expression by timepoint. Also filtering for a minimum gene count of 10, such that ~12% of the samples must have a gene count of >10 in order for the gene to remain in the data set.  
```{r}
filt <- filterfun(pOverA(0.12,10))

#create filter for the counts data
gfilt <- genefilter(counts, filt)

#identify genes to keep by count filter
gkeep <- counts[gfilt,]

#identify gene lists
gn.keep <- rownames(gkeep)

#gene count data filtered in PoverA, P percent of the samples have counts over A
counts_filt <- as.data.frame(counts[which(rownames(counts) %in% gn.keep),])

#How many rows do we have before and after filtering?
nrow(counts) #Before
nrow(counts_filt) #After

# Save filtered counts matrix 
write.csv(counts_filt, file = "../output/Molecular/mRNA_polyA/mRNA_polyA_filtered_counts.csv")
```

Before filtering, there were 27366 genes; after filtering, there were 9663 genes. 

Read in metadata 
```{r}
meta <- read.csv("../data/sample_metadata/tube_meta_mcap.csv") %>%
  dplyr::select(TubeID, Date, hpf, Metric, EmbryosPerTube) %>%
  filter(Metric == "Molecular")

# Filter for sequenced samples
meta_filt <- meta %>%
  filter(TubeID %in% colnames(counts_filt))
```

Remove M14 and M61 - these are outliers (based on PCA below)
```{r}
counts_filt <- counts_filt[, !(colnames(counts_filt) %in% c("M61"))]
meta_filt <- meta_filt[!(meta_filt$TubeID %in% c("M61")), ]
```

Reorder columns in filtered counts matrix so that it matches the metadata
```{r}
counts_filt <- counts_filt[, meta_filt$TubeID]

# Sanity check! 
meta_filt$TubeID %in% colnames(counts_filt) #are all of the sample names (rows) in the metadata df in the gene count matrix? Should be TRUE. 
all(rownames(meta_filt$TubeID) == colnames(counts_filt)) #are they the same in the same order? Should be TRUE
```

Set levels of factors 
```{r}
meta_filt$hpf <- factor(meta_filt$hpf, levels = c("1 hpf", "4 hpf", "9 hpf", "14 hpf", "22 hpf", "28 hpf", "48 hpf", "72 hpf"))
```

Construct DESeq data set
```{r}
gdds <- DESeqDataSetFromMatrix(countData = counts_filt, colData = meta_filt, design = ~hpf)
```

Log transform data using VST for visualization purposes, which will deal with low count variability by calculating within-group variation (if blind=FALSE). It doesn't remove design from the data so we can see how all samples look (check if grouping by lifestage, batch effects, etc). To do this, calculate size factors of samples,which is a rough estimate of how many reads each sample contains compared to others. To use VST, size factors should be <4.
```{r}
SF.gdds <- estimateSizeFactors(gdds) #estimate size factors to determine if we can use vst  to transform our data. Size factors should be less than 4 for us to use vst
print(sizeFactors(SF.gdds)) #View size factors

all(sizeFactors(SF.gdds)) < 4 # should return TRUE
```

All size factors are <4, so we can use vst. 
```{r}
gvst <- vst(gdds, blind=FALSE) #apply a variance stabilizing transformation to minimize effects of small counts and normalize wrt library size
head(assay(gvst), 3) #view transformed gene count data for the first three genes in the dataset.  
```

Plot a heatmap to sample to sample distances  
```{r}
gsampleDists <- dist(t(assay(gvst))) #calculate distance matix
gsampleDistMatrix <- as.matrix(gsampleDists) #distance matrix
rownames(gsampleDistMatrix) <- colnames(gvst) #assign row names
colnames(gsampleDistMatrix) <- NULL #assign col names
#colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255) #assign colors

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}

pht<-pheatmap(gsampleDistMatrix, #plot matrix
         clustering_distance_rows=gsampleDists, #cluster rows
         clustering_distance_cols=gsampleDists) #cluster columns
```

Generate a PCA of all genes    
```{r}
pcaData <- plotPCA(gvst, intgroup = c("hpf"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color = hpf, label = name)) + 
  geom_point(size = 3) + 
  geom_text(vjust = 1.5, hjust = 1.5) + # This adds the labels
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed()
```

PCA separates beautifully!! There are a couple of outliers -- M61 from 28 hpf and M72 from 48 hpf. Will need to remove these before proceeding. 

Generate a publication style PCA  
```{r}
pcaData <- plotPCA(gvst, intgroup = c("hpf"), returnData = TRUE)

percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color = hpf, label = name)) + 
  #geom_point(size = 3) + 
  #geom_jitter(size = 4, width = 1.5, height = 1.5, stroke = 1) +
  geom_jitter(aes(fill = hpf), size = 5, width = 1.5, height = 1.5, shape = 21, color = "black", stroke = 0.5) +
  #geom_text(vjust = 1.5, hjust = 1.5) + # This adds the labels
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  scale_color_discrete() +
  coord_fixed() +
  theme_linedraw() +
  theme(legend.text = element_text(size = 13),
        plot.background = element_blank(),
        legend.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 15, face = "bold"),
        #axis.title.y = element_text(vjust = -1.5),
        plot.title = element_text(size = 20, face = "italic", hjust = 0.5))

# Save 
ggsave("../output/Molecular/mRNA_polyA/PCA_all_genes_polyA.pdf", plot = last_plot())
ggsave("../output/Molecular/mRNA_polyA/PCA_all_genes_polyA.png", plot = last_plot())
```

Plot gene count for specific genes of interest (or all genes) over time (vst transformed counts)
```{r}
# Transform counts
vst_counts <- vst(gdds, blind = TRUE)
vst_mat <- assay(vst_counts)

# Pivot to long
vst_long <- vst_mat %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  pivot_longer(-gene, names_to = "TubeID", values_to = "vst_count")

# Join with metadata
vst_long <- vst_long %>%
  left_join(meta_filt, by = "TubeID")

# Plot
gene_of_interest <- "Montipora_capitata_HIv3___RNAseq.g41714.t1"
plot_summary <- vst_long %>%
  filter(gene == gene_of_interest) %>%
  group_by(hpf) %>%
  summarize(
    mean_vst = mean(vst_count, na.rm = TRUE),
    se_vst = sd(vst_count, na.rm = TRUE) / sqrt(n())) %>%
  mutate(hpf = factor(hpf, levels = unique(meta_filt$hpf)))  # consistent ordering
ggplot(plot_summary, aes(x = hpf, y = mean_vst, group = 1)) +
  geom_point(size = 3, color = "darkblue") +
  geom_line(color = "darkblue", linewidth = 0.5) +
  geom_errorbar(aes(ymin = mean_vst - se_vst, ymax = mean_vst + se_vst), width = 0.2, color = "darkblue") +
  labs(x = "Hours Post Fertilization (hpf)",
    y = "Mean VST Normalized Count ± SE"
  ) +
  theme_minimal()
```

Run test using the LRT to look for effect of lifestage as a whole across our experiment. There is only one main effect, so the reduced model is just the intercept. We are using LRT per recommendations from the DESeq2 [manual](https://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html), which states about time-series experiments: "There are a number of ways to analyze time-series experiments, depending on the biological question of interest. In order to test for any differences over multiple time points, once can use a design including the time factor, and then test using the likelihood ratio test as described in the following section, where the time factor is removed in the reduced formula."

I initially ran DESeq2 with LRT but the results (number of DEGs and volcano/MA plots) didn't look correct. Now, I am running it with Wald!
```{r, message = FALSE}
DEG <- DESeq(gdds)
```

Save all DE results, regardless of significance 
```{r}
DEG_results_all<-as.data.frame(results(DEG))
DEG_results_all$gene <- rownames(DEG_results_all)
rownames(DEG_results_all) <- NULL
dim(DEG_results_all)

# Save as csv 
write.csv(DEG_results_all, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_wald_polyA.csv")
#DEG_results_all <- read.csv("../output/Molecular/mRNA_polyA/mRNA_DE_results_all_wald_polyA.csv")
```

Filter by adjusted p value <0.05.
```{r}
DEG_results <- subset(DEG_results_all, padj<0.05)
dim(DEG_results)

# Save as csv 
write.csv(DEG_results, "../output/Molecular/mRNA_polyA/mRNA_DE_results_wald_polyA.csv")
```

Conduct VST transformation for visualization.  
```{r}
DEG_results_vst <- gdds[unique(DEG_results$gene)]
dim(DEG_results_vst)

DEG_results_vst <- varianceStabilizingTransformation(DEG_results_vst) 
```

Plot a heatmap to sample to sample distances  
```{r}
gsampleDists <- dist(t(assay(DEG_results_vst))) #calculate distance matix
gsampleDistMatrix <- as.matrix(gsampleDists) #distance matrix
rownames(gsampleDistMatrix) <- colnames(DEG_results_vst) #assign row names
colnames(gsampleDistMatrix) <- NULL #assign col names
#colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255) #assign colors

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}

pht<-pheatmap(gsampleDistMatrix, #plot matrix
         clustering_distance_rows=gsampleDists, #cluster rows
         clustering_distance_cols=gsampleDists) #cluster columns
```

Generate a PCA of DEGs    
```{r}
pcaData_DEG <- plotPCA(DEG_results_vst, intgroup = c("hpf"), returnData = TRUE)

percentVar_DEG <- round(100 * attr(pcaData_DEG, "percentVar"))
ggplot(pcaData_DEG, aes(PC1, PC2, color = hpf, label = name)) + 
  #geom_point(size = 3) + 
  #geom_jitter(size = 4, width = 1.5, height = 1.5, stroke = 1) +
  geom_jitter(aes(fill = hpf), size = 5, width = 1.5, height = 1.5, shape = 21, color = "black", stroke = 0.5) +
  #geom_text(vjust = 1.5, hjust = 1.5) + # This adds the labels
  xlab(paste0("PC1: ", percentVar_DEG[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar_DEG[2], "% variance")) +
  scale_color_discrete() +
  coord_fixed() +
  theme_linedraw() +
  theme(legend.text = element_text(size = 13),
        plot.background = element_blank(),
        legend.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 15, face = "bold"),
        #axis.title.y = element_text(vjust = -1.5),
        plot.title = element_text(size = 20, face = "italic", hjust = 0.5))
```

Look at specific contrasts 
```{r}
resultsNames(DEG)
#"hpf_4.hpf_vs_1.hpf"  "hpf_9.hpf_vs_1.hpf"  "hpf_14.hpf_vs_1.hpf"
# "hpf_22.hpf_vs_1.hpf" "hpf_28.hpf_vs_1.hpf" "hpf_48.hpf_vs_1.hpf" "hpf_72.hpf_vs_1.hpf"'
```

1 v 4 hpf 
```{r}
# Wald
results_1v4_wald <- as.data.frame(results(DEG, name = c("hpf_4.hpf_vs_1.hpf")))
results_1v4_wald <- results_1v4_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 4 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v4_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v4_wald_polyA.csv")

results_1v4_wald_sig <- results_1v4_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v4_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v4_wald_sig$log2FoldChange > 0) # upregulated in 4hpf
write.csv(results_1v4_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v4_wald_polyA.csv") 

ggplot(data = results_1v4_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 4 hpf")) +
  theme_cowplot()
```

1 v 9 hpf 
```{r}
# Wald
results_1v9_wald <- as.data.frame(results(DEG, name = c("hpf_9.hpf_vs_1.hpf")))
results_1v9_wald <- results_1v9_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 9 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v9_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v9_wald_polyA.csv")

results_1v9_wald_sig <- results_1v9_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v9_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v9_wald_sig$log2FoldChange > 0) # upregulated in 9hpf
write.csv(results_1v9_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v9_wald_polyA.csv") 

ggplot(data = results_1v9_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 9 hpf")) +
  theme_cowplot()
```

1 v 14 hpf 
```{r}
# Wald
results_1v14_wald <- as.data.frame(results(DEG, name = c("hpf_14.hpf_vs_1.hpf")))
results_1v14_wald <- results_1v14_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 14 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v14_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v14_wald_polyA.csv")

results_1v14_wald_sig <- results_1v14_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v14_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v14_wald_sig$log2FoldChange > 0) # upregulated in 14hpf
write.csv(results_1v14_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v14_wald_polyA.csv") 

ggplot(data = results_1v14_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 14 hpf")) +
  theme_cowplot()
```

1 v 22 hpf 
```{r}
# Wald
results_1v22_wald <- as.data.frame(results(DEG, name = c("hpf_22.hpf_vs_1.hpf")))
results_1v22_wald <- results_1v22_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 22 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v22_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v22_wald_polyA.csv")

results_1v22_wald_sig <- results_1v22_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v22_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v22_wald_sig$log2FoldChange > 0) # upregulated in 22hpf
write.csv(results_1v22_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v22_wald_polyA.csv") 

ggplot(data = results_1v22_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 22 hpf")) +
  theme_cowplot()
```

1 v 28 hpf 
```{r}
# Wald
results_1v28_wald <- as.data.frame(results(DEG, name = c("hpf_28.hpf_vs_1.hpf")))
results_1v28_wald <- results_1v28_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 28 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v28_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v28_wald_polyA.csv")

results_1v28_wald_sig <- results_1v28_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v28_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v28_wald_sig$log2FoldChange > 0) # upregulated in 28hpf
write.csv(results_1v28_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v28_wald_polyA.csv") 

ggplot(data = results_1v28_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 28 hpf")) +
  theme_cowplot()
```

1 v 48 hpf 
```{r}
# Wald
results_1v48_wald <- as.data.frame(results(DEG, name = c("hpf_48.hpf_vs_1.hpf")))
results_1v48_wald <- results_1v48_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 48 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v48_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v48_wald_polyA.csv")

results_1v48_wald_sig <- results_1v48_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v48_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v48_wald_sig$log2FoldChange > 0) # upregulated in 48hpf
write.csv(results_1v48_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v48_wald_polyA.csv") 

ggplot(data = results_1v48_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 48 hpf")) +
  theme_cowplot()
```

1 v 72 hpf 
```{r}
# Wald
results_1v72_wald <- as.data.frame(results(DEG, name = c("hpf_72.hpf_vs_1.hpf")))
results_1v72_wald <- results_1v72_wald %>%
  mutate(
    regulation = case_when(
      padj < 0.05 & log2FoldChange < -1 ~ "Up at 1 hpf",
      padj < 0.05 & log2FoldChange > 1 ~ "Up at 72 hpf",
      TRUE ~ "Not significant"))
write.csv(results_1v72_wald, "../output/Molecular/mRNA_polyA/mRNA_DE_results_all_1v72_wald_polyA.csv")

results_1v72_wald_sig <- results_1v72_wald %>%
  filter(padj < 0.05 & abs(log2FoldChange) >= 1)
sum(results_1v72_wald_sig$log2FoldChange < 0) # upregulated in 1hpf
sum(results_1v72_wald_sig$log2FoldChange > 0) # upregulated in 72hpf
write.csv(results_1v72_wald_sig, "../output/Molecular/mRNA_polyA/mRNA_DE_results_sig_1v72_wald_polyA.csv") 

ggplot(data = results_1v72_wald, aes(x = log2FoldChange, y = -log10(padj), color = regulation)) +
  geom_vline(xintercept = c(-1, 1), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
  geom_point() +
  scale_color_manual(values = c("grey", "#00AFBB", "purple"),
                     labels = c("Not significant", "Upregulated at 1 hpf", "Upregulated at 72 hpf")) +
  theme_cowplot()
```

Get DESeq2 comparisons across time point
```{r}
timepoints <- c(4, 9, 14, 22, 28, 48, 72) # hpf timepoints
result_list <- list()

for (tp in timepoints) {
  res <- results(DEG, name = paste0("hpf_", tp, ".hpf_vs_1.hpf"))
  res_df <- as.data.frame(res)
  res_df$gene <- rownames(res_df)
  res_df$timepoint <- tp
  result_list[[as.character(tp)]] <- res_df
}

# Combine into one dataframe
all_results <- bind_rows(result_list)
```

Find earliest time point with >2 LFC 
```{r}
# Filter for genes with >2-fold decrease (log2FC < -1)
downregulated <- all_results %>%
  filter(log2FoldChange < -1)

# For each gene, find the earliest timepoint
clearance_time_rRNA <- downregulated %>%
  group_by(gene) %>%
  summarize(clearance_time_rRNA = min(timepoint))
```

Plot 
```{r}
# Count how many genes per clearance time
clearance_counts_rRNA <- clearance_time_rRNA %>%
  group_by(clearance_time_rRNA) %>%
  summarise(n_genes = n()) %>%
  ungroup()

# Bar plot
ggplot(clearance_counts_rRNA, aes(x = factor(clearance_time_rRNA), y = n_genes)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    x = "Clearance Time (hpf, polyA)",
    y = "Number of genes") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  )

# Dot plot
ggplot(clearance_counts_rRNA, aes(x = clearance_time_rRNA, y = n_genes)) +
  geom_point(size = 5, color = "dodgerblue") +
  geom_line(group = 1, color = "dodgerblue") +
  theme_minimal() +
  labs(
    x = "Clearance Time (hpf, polyA)",
    y = "Number of genes"
  )
```
