---
title: "Physiology"
output: html_document
date: "2023-09-26"
---

This script analyzes and plots data for Developmental Timeseries 2023 physiology measurements including protein, carbohydrates, and lipids, and correlations between these variables. This script is based off of AH's [script](https://github.com/AHuffmyer/SymbioticIntegration/blob/main/Mcap2021/Scripts/Physiology.Rmd).

# Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(tidyverse)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)
library(ggsignif)
library(ggstatsplot)


getwd()
```

# Set plot theme 

```{r}
custom_theme<-theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        legend.text=element_text(size=12, color="black"), 
        legend.title=element_text(size=12, face="bold", color="black"))
```

# Protein

## Load data 

Read in datafiles

```{r}
protein <- read.csv("../2023/data/Physiology/SolubleProtein_Data.csv")

prot_meta <- read.csv("../2023/data/Physiology/SolubleProtein_Meta.csv")
prot_meta <- prot_meta %>%
  mutate(Homo_vol = ifelse(Sample.Type == "Sample", 25, NA)) %>%
  mutate(Resusp_vol = ifelse(Sample.Type == "Sample", 500, NA)) 

meta <- read.csv("../2023/data/sample_metadata/tube_meta.csv")
```

## Prepare dataframes 

Remove unneeded rows from meta dataframe 

```{r}
meta <- meta %>%
  filter(Metric == "Physiology")
```

Merge Files and rename columns

```{r}
protein <- merge(prot_meta, protein, by = c("Well", "Run"))
protein <- left_join(protein, meta, by=c("TubeID"))

# Blank correction for each run separately

Blank <- protein %>% 
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(x562))

protein$abs.corr <- protein$x562 - Blank$blk.avg
```

Plot regressions for each run.  

```{r}
protein %>% 
  filter(Sample.Type == "Standard")%>%

  ggplot(aes(x=Concentration, y=abs.corr))+
  facet_wrap(~ Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  ggtitle("Run")+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = after_stat(eq.label))) +
  custom_theme
```

Run a loop to extract regression for each run and calculate protein concentration.  

```{r}
run_list<-c("1", "2", "3", "4", "5", "6", "7", "8")

df_protein = data.frame()

for (run in run_list) {

  #subset data
Standard <- protein %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==run)
  
#generate standard curve equation
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 

#select samples
Samples <- protein %>% #subsetting Samples
  filter(Sample.Type == "Sample") %>%
  filter(Run==run)

#calculate concentration
Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration

#add run column
#Samples$Run<-run

#normalize to homogenate volume
Samples$Protein.ug <- Samples$Concentration * (Samples$Resusp_vol / 1000) * (25/Samples$Homo_vol)

#join dataframes
df <- data.frame(Samples)
df_protein <- rbind(df_protein,df)

} 
```


Plot results 

```{r}
# Keep only Mcap results
df_protein_mcap <- df_protein %>%
  filter(Species == "Montipora capitata")

# Order the 'hpf' variable in a specific order
df_protein_mcap$hpf <- factor(df_protein_mcap$hpf, levels = c("1 hpf", "4 hpf", "9 hpf", "14 hpf", "22 hpf", "28 hpf", "48 hpf", "72 hpf"))

Protein.Plot <- df_protein_mcap%>%
  ggplot(aes(x=hpf, y=Protein.ug, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  scale_fill_manual(values = c("blue", "red", "gray")) +
  custom_theme; Protein.Plot
```

## Calculate protein/ug/larvae 

Calculate protein/ug/larvae
```{r}
df_protein_mcap<-df_protein_mcap%>%
  dplyr::select(TubeID, Treatment, Fraction, Protein.ug, hpf, EmbryosPerTube)%>%
  group_by(Fraction, Treatment, hpf)%>%
  mutate(Protein.ug.larvae = Protein.ug/EmbryosPerTube)%>%
  na.omit()
  #mutate(Protein.mg = Protein.ug/1000)%>%
  #mutate(Protein.mg.larvae = Protein.mg/EmbryosPerTube)
  #dplyr::summarise(mean_protein=mean(Protein.ug.larvae))



# %>%
#   spread(Fraction, mean_protein)%>%
#   mutate(ratio=Host/Holobiont)%>%
#   mutate(sym_holo_ratio=Symbiont/Holobiont)%>%
#   na.omit()
```

### Plot data 

Plot holo, host and sym protein 
```{r}
holo.protein.plot <- df_protein_mcap%>%
  filter(Fraction == "Holobiont")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; holo.protein.plot

host.protein.plot <- df_protein_mcap%>%
  filter(Fraction == "Host")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; host.protein.plot

sym.protein.plot <- df_protein_mcap%>%
  filter(Fraction == "Symbiont")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; sym.protein.plot
```

Remove outliers 
```{r}
remove_outliers <- function(x) {
  qnt <- quantile(x, probs=c(.25, .75))
  H <- 1.5 * IQR(x)
  x[x > (qnt[2] + H) | x < (qnt[1] - H)] <- NA
  return(x)
}

df_protein_mcap_clean <- df_protein_mcap %>%
  group_by(Treatment, Fraction, hpf) %>%
  mutate(Protein.ug.larvae = remove_outliers(Protein.ug.larvae)) %>%
  ungroup() %>%
  na.omit(Protein.ug.larvae)
```

Plot holo, host and sym protein with outliers removed 
```{r}
holo.protein.plot <- df_protein_mcap_clean%>%
  filter(Fraction == "Holobiont")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; holo.protein.plot

host.protein.plot <- df_protein_mcap_clean%>%
  filter(Fraction == "Host")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; host.protein.plot

sym.protein.plot <- df_protein_mcap_clean%>%
  filter(Fraction == "Symbiont")%>%
  ggplot(aes(x=hpf, y=Protein.ug.larvae, fill = Treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("hpf") + 
  #facet_grid(. ~ Fraction) +  # Use facet_grid to separate bars by 'Fraction'
  #ylab(expression(bold("Soluble Protein " (ug)))) + #Axis titles
  scale_fill_manual(values = c("blue", "red")) +
  custom_theme; sym.protein.plot
```

```{r}
sym <- df_protein_mcap_clean %>%
  filter(Fraction == "Symbiont")

res.aov <- aov(Protein.ug.larvae ~ Treatment*hpf, data = sym)
summary(res.aov)

holo <- df_protein_mcap_clean %>%
  filter(Fraction == "Holobiont")

res.aov <- aov(Protein.ug.larvae ~ Treatment*hpf, data = holo)
summary(res.aov)

host <- df_protein_mcap_clean %>%
  filter(Fraction == "Host")

res.aov <- aov(Protein.ug.larvae ~ Treatment*hpf, data = host)
summary(res.aov)

TukeyHSD(res.aov)
```

















