---
title: "GO enrichment"
author: "Jill Ashey"
date: "2024-12-28"
output: html_document
---

GO enrichment of Mcap miRNA targets that are decreasing as development progresses.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#devtools::install_github("stemangiola/tidyHeatmap")

library(tidyverse)
library(topGO)
library(tidyHeatmap)
library(purrr)
library(scales)
library(igraph)
library(tidygraph)
library(ggraph)
```

Read in correlation data generated from mirTarRnaSeq 
```{r}
corr <- read.csv("../output/Molecular/miRNA_mRNA_interactions/mirtarrnaseq_correlations_data.csv")
head(corr)
```

Make df with only gene and miRNA info 
```{r}
corr_gene_mirna <- corr %>%
  dplyr::select(V1, V2) %>%
  dplyr::rename(mirna = V1, gene = V2) 
```

Read in annotation file and make tab file for topGO
```{r}
annot <- read.delim("~/Desktop/GFFs/mcap/V3/Montipora_capitata_HIv3.genes.EggNog_results.txt")%>%
  dplyr::select(X.query, GOs) %>%
  mutate(across(everything(), ~ na_if(., "-")))

# Make sure all GO terms are separated by ;
annot$GOs <- gsub(",", ";", annot$GOs)

#write_tsv(annot, file = "../data/Molecular/Mcap_V3_gene2go.tab")
```

Make list of genes for input to topGO
```{r}
# Genes of interest - ie those targeted by miRNAs 
clust_genes <- as.character(corr_gene_mirna$gene)

# All genes 
all_genes <- as.character(annot$X.query)

# Apply 1 or 0 if gene is gene of interest 
GeneList <- factor(as.integer(all_genes %in% clust_genes))
names(GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test to assess whether specific GO terms are overrepresented in the genes targeted by miRNAs. 

Read in gene-to-go-mappings
```{r}
gene2go_topgo <- readMappings("../data/Molecular/Mcap_V3_gene2go.tab", IDsep=";")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_BP <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_BP_FE <- runTest(GO_BP, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_BP_En <- GenTable(GO_BP, Fisher = GO_BP_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_BP_En$Fisher<-as.numeric(GO_BP_En$Fisher)
GO_BP_En_sig<-GO_BP_En[GO_BP_En$Fisher<0.05,]
```

Merge `GO_BP_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
annot <- annot %>%
  separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
annot$GOs <- trimws(annot$GOs)
GO_BP_En_sig$GO.ID <- trimws(GO_BP_En_sig$GO.ID)

# Join the datasets based on GO term
GO_BP_En_sig_gene <- annot %>%
  left_join(GO_BP_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_BP_En_sig_gene$ontology <- "Biological Processes"
```

Merge with corr and miRNA data
```{r}
GO_BP_En_sig_gene_mirna <- GO_BP_En_sig_gene %>%
  full_join(corr, by = c("X.query" = "V2")) %>%
  na.omit()

length(unique(GO_BP_En_sig_gene_mirna$X.query))
length(unique(GO_BP_En_sig_gene_mirna$V1))
length(unique(GO_BP_En_sig_gene_mirna$GOs))
```

### Cellular Components 

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_CC <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_CC_FE <- runTest(GO_CC, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_CC_En <- GenTable(GO_CC, Fisher = GO_CC_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_CC_En$Fisher<-as.numeric(GO_CC_En$Fisher)
GO_CC_En_sig<-GO_CC_En[GO_CC_En$Fisher<0.05,]
```

Merge `GO_CC_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
# annot <- annot %>%
#   separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
# annot$GOs <- trimws(annot$GOs)
GO_CC_En_sig$GO.ID <- trimws(GO_CC_En_sig$GO.ID)

# Join the datasets based on GO term
GO_CC_En_sig_gene <- annot %>%
  left_join(GO_CC_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_CC_En_sig_gene$ontology <- "Cellular Components"
```

Merge with corr and miRNA data
```{r}
GO_CC_En_sig_gene_mirna <- GO_CC_En_sig_gene %>%
  full_join(corr, by = c("X.query" = "V2")) %>%
  na.omit()

length(unique(GO_CC_En_sig_gene_mirna$X.query))
length(unique(GO_CC_En_sig_gene_mirna$V1))
length(unique(GO_CC_En_sig_gene_mirna$GOs))
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_MF <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```
Run GO enrichment test 
```{r}
GO_MF_FE <- runTest(GO_MF, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_MF_En <- GenTable(GO_MF, Fisher = GO_MF_FE, orderBy = "Fisher", numChar = 51, topNodes = 100)
```

Filter by significant results
```{r}
GO_MF_En$Fisher<-as.numeric(GO_MF_En$Fisher)
GO_MF_En_sig<-GO_MF_En[GO_MF_En$Fisher<0.05,]
```

Merge `GO_MF_En_sig` with GO and gene info. 
```{r}
# Separate GO terms 
# annot <- annot %>%
#   separate_rows(GOs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
# annot$GOs <- trimws(annot$GOs)
GO_MF_En_sig$GO.ID <- trimws(GO_MF_En_sig$GO.ID)

# Join the datasets based on GO term
GO_MF_En_sig_gene <- annot %>%
  left_join(GO_MF_En_sig, by = c("GOs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_MF_En_sig_gene$ontology <- "Molecular Functions"
```

Merge with corr and miRNA data
```{r}
GO_MF_En_sig_gene_mirna <- GO_MF_En_sig_gene %>%
  full_join(corr, by = c("X.query" = "V2")) %>%
  na.omit()

length(unique(GO_MF_En_sig_gene_mirna$X.query))
length(unique(GO_MF_En_sig_gene_mirna$V1))
length(unique(GO_MF_En_sig_gene_mirna$GOs))
```

### Join ontologies 

Bind so there is a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_En_sig_gene_mirna <- rbind(GO_BP_En_sig_gene_mirna, GO_CC_En_sig_gene_mirna, GO_MF_En_sig_gene_mirna)

# Calculate proportion of significant v annotated genes 
GO_En_sig_gene_mirna <- GO_En_sig_gene_mirna %>%
  mutate(sig.prop = Significant/Annotated) %>%
  na.omit()

# Save as csv 
#write.csv(GO_En_sig_gene_mirna, "../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets.csv")
```

Plot BP and MF only 
```{r}
GO_En_sig_gene_mirna_bp <- GO_En_sig_gene_mirna %>%
  dplyr::filter(ontology == "Biological Processes") #%>%
  #dplyr::filter(V1 == "Montipora_capitata_HIv3___Scaffold_8_586427")

bp_plot<-ggplot(GO_En_sig_gene_mirna_bp, aes(x = Term, y = Fisher, fill = Fisher, size = sig.prop)) +
  #expand_limits(y = 1.5) +
  #ylim(1, 7.25) +
  # Add horizontal lines with a single aesthetic value
  #geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  #geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  #scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); bp_plot
#ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_BP.pdf", plot = bp_plot, height = 25, width = 20)
#ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_BP.png", plot = bp_plot, height = 25, width = 20)

GO_En_sig_gene_mirna_mf <- GO_En_sig_gene_mirna %>%
  dplyr::filter(ontology == "Molecular Functions") #%>%
  #dplyr::filter(V1 == "Montipora_capitata_HIv3___Scaffold_8_586427")

mf_plot<-ggplot(GO_En_sig_gene_mirna_mf, aes(x = Term, y = Fisher, fill = Fisher, size = sig.prop)) +
  #expand_limits(y = 1.5) +
  #ylim(1, 7.25) +
  # Add horizontal lines with a single aesthetic value
  #geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  #geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  #scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); mf_plot
#ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_MF.pdf", plot = mf_plot, height = 25, width = 20)
#ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_MF.png", plot = mf_plot, height = 25, width = 20)
```

Create network plot that shows the miRNAs that regulate mRNAs 
```{r}
# Read in data if needed 
#blah <- read.csv("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets.csv")

annot <- read.delim("~/Desktop/GFFs/mcap/V3/Montipora_capitata_HIv3.genes.EggNog_results.txt")

test <- annot %>%
  full_join(GO_En_sig_gene_mirna, by = "X.query") %>%
  na.omit() %>%
  dplyr::select(X.query, Description, Term, Preferred_name, V1, value) #%>%
  #dplyr::filter(V1 == "Montipora_capitata_HIv3___Scaffold_3_149822") #%>%
  #dplyr::filter(V1 %in% c("Montipora_capitata_HIv3___Scaffold_10_729473", "Montipora_capitata_HIv3___Scaffold_151_1075760", "Montipora_capitata_HIv3___Scaffold_11_826703")) # add this if you want to look at a specific miRNA of interest; otherwise, comment out 

# Count the number of target mRNAs for each miRNA
result <- test %>%
  group_by(V1) %>%
  summarise(num_queries = n_distinct(X.query))

# Plot number of targets for each miRNA
target_num <- ggplot(result, aes(x = reorder(V1, -num_queries), y = num_queries)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "miRNA", y = "Number of Targets", 
       title = "Number of Targets per miRNA") +
  coord_flip(); target_num

# Create the graph from the data frame
g <- graph_from_data_frame(test[, c("V1", "Description", "value")], directed = FALSE)

# Add edge attributes
E(g)$weight <- abs(E(g)$value)  # Use absolute PCC for edge weight
E(g)$color <- ifelse(E(g)$value > 0, "blue", "red")  # Blue for positive, red for negative correlations - in this case, we only have negative correlations 

# Add node attributes
V(g)$type <- ifelse(V(g)$name %in% test$V1, "miRNA", "mRNA")

# Convert to tbl_graph for ggraph
g_tbl <- as_tbl_graph(g)

# Create plot
network_plot <- ggraph(g_tbl, layout = "nicely") +
  geom_edge_link(aes(edge_width = weight, color = color)) +
  geom_node_point(aes(color = type), size = 7) +
  #geom_node_text(aes(label = name), repel = TRUE, size = 10, 
  #               family = "Arial", fontface = "bold") +
  scale_edge_width(range = c(1, 3), name = "Correlation Strength") +
  scale_color_manual(values = c("miRNA" = "purple", "mRNA" = "orange", 
                                "Positive correlation" = "#3C5488", 
                                "Negative correlation" = "#DC0000"),
                     name = "Node/Edge Type") +
  theme_graph(base_family = "Arial") +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.caption = element_text(size = 8, hjust = 1)
  ); network_plot

#ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_network.pdf", plot = network_plot, height = 49, width = 40)
ggsave("../output/Molecular/miRNA_mRNA_interactions/GO_en_sig_miRNA_targets_network.png", plot = network_plot, height = 20, width = 15)
```

If the network plot does not plot due to memory limits, restart R and plot again.

Next step: plot expression over time for miRNAs and their target mRNAs 

ONLY UP TO 22 HPF

Read in mRNA count data
```{r}
mrna_counts <- read.csv("../output/Molecular/mRNA/mRNA_filtered_counts.csv")

# Select specific columns 
#mrna_counts <- mrna_counts %>%
#  dplyr::select(X, M9, M13, M23, M35, M52, M60, M72, M85)
mrna_counts <- mrna_counts %>%
  dplyr::select(X, M9, M13, M23, M35, M52)

# Rename cols based on hpf 
#colnames(mrna_counts) <- c("gene", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf", "28_hpf", "48_hpf", "72_hpf")
colnames(mrna_counts) <- c("gene", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf")
```

Read in miRNA count data 
```{r}
mirna_counts <- read.delim("../output/Molecular/smRNA/mirdeep2/miRNAs_expressed_all_samples_1733514422.csv", header = T)
mirna_counts <- as.data.frame(mirna_counts)

# Select specific columns 
#mirna_counts <- mirna_counts %>%
#  dplyr::select(X.miRNA, s09.norm., s13.norm., s23.norm., s35.norm., s52.norm., s60.norm., s72.norm., s85.norm.)
mirna_counts <- mirna_counts %>%
  dplyr::select(X.miRNA, s09.norm., s13.norm., s23.norm., s35.norm., s52.norm.)

# Rename cols based on hpf 
colnames(mirna_counts) <- c("mirna", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf")
```

Extract this info from counts dfs
```{r}
mirna_data <- mirna_counts %>%
  filter(mirna %in% GO_En_sig_gene_mirna$V1) %>%
  pivot_longer(cols = -mirna, names_to = "hpf", values_to = "count") %>%
  mutate(type = "miRNA") %>%
  distinct(mirna, hpf, .keep_all = TRUE)  # Remove any duplicates

mrna_data <- mrna_counts %>%
  filter(gene %in% GO_En_sig_gene_mirna$X.query) %>%
  pivot_longer(cols = -gene, names_to = "hpf", values_to = "count") %>%
  mutate(type = "mRNA") %>%
  distinct(gene, hpf, .keep_all = TRUE)  # Remove any duplicates
```

Plot the miRNAs with their interacting mRNAs over time 
```{r}
create_mirna_plot <- function(specific_mirna) {
  # Find interacting mRNAs
  interacting_mrnas <- corr %>%
    dplyr::filter(V1 == specific_mirna) %>%
    dplyr::pull(V2)
  
  # Extract count data
  mirna_counts <- mirna_data %>%
    dplyr::filter(mirna == specific_mirna) %>%
    dplyr::select(hpf, count, type)
  
  mrna_counts <- mrna_data %>%
    dplyr::filter(gene %in% interacting_mrnas) %>%
    dplyr::select(hpf, count, type, gene)
  
  # Combine data
  combined_data <- bind_rows(
    mirna_counts,
    mrna_counts
  ) %>%
    mutate(hpf = as.numeric(gsub("_hpf", "", hpf)))
  
  # Create the plot
  ggplot(combined_data, aes(x = hpf, y = count, color = type, group = interaction(type, gene))) +
    geom_line(size = 1) +
    geom_point(size = 3) +
    scale_x_continuous(breaks = unique(combined_data$hpf)) +
    scale_y_log10(labels = scales::comma) +
    scale_color_manual(values = c("miRNA" = "purple", "mRNA" = "orange")) +
    labs(title = paste("Expression of", specific_mirna, "and its interacting mRNAs"),
         subtitle = "During early development",
         x = "Hours post fertilization",
         y = "Count (log scale)",
         color = "RNA type") +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title = element_text(face = "bold", size = 12),
      axis.text = element_text(size = 10),
      legend.title = element_text(face = "bold", size = 12),
      legend.text = element_text(size = 10),
      legend.position = "right",
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, size = 1)
    )
}

# Generate plots for all miRNAs
all_mirnas <- unique(mirna_data$mirna)
all_plots <- map(all_mirnas, create_mirna_plot)

# Save plots to a PDF file
# pdf("../output/Molecular/miRNA_mRNA_interactions/all_mirna_plots_until_22hpf.pdf", width = 10, height = 8)
# for(plot in all_plots) {
#   print(plot)
# }
# dev.off()

```



