---
title: "TPM DEGs"
author: "Jill Ashey"
date: "2025-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read in GFF to calculate transcript length
```{r}
gff <- read.delim("~/Desktop/GFFs/mcap/V3/Montipora_capitata_HIv3.genes.gff3", header = F)
colnames(gff) <- c("seqid", "source", "type", "start", "end", "score", "strand", "phase", "attributes")
```

Select only transcripts and calculate transcript length
```{r}
gff_transcript <- gff %>%
  filter(type == "transcript") %>%
  mutate(length = end - start)
```

Save gene id and length as csv
```{r}
gff_transcript$gene_id <- gsub("ID=", "", gff_transcript$attributes)

lengths_df <- gff_transcript[, c("length", "gene_id")]
write.csv(lengths_df, "../data/Molecular/transcript_lengths.csv", row.names = FALSE)
```

### rRNA depleted samples 

Read in filtered counts data
```{r}
counts <- read.csv("../output/Molecular/mRNA/mRNA_filtered_counts.csv")
colnames(counts)[1] <- "gene_id"

# Remove M14 and M61, they are outliers 
counts <- counts[, !(colnames(counts) %in% c("M14", "M61"))]
```

Join counts with length info. This is needed to calculate TPM. 
```{r}
counts_length <- counts %>%
  inner_join(lengths_df, by = "gene_id")
```

Calculate TPM
```{r}
# Get gene lengths in kilobases
length_kb <- counts_length$length / 1000

# Extract count columns (excluding gene_id and length)
count_cols <- setdiff(names(counts_length), c("gene_id", "length"))

# Calculate RPK (Reads Per Kilobase)
RPK <- sweep(counts_length[, count_cols], 1, length_kb, FUN = "/")

# Sum all RPK values for each sample
sum_RPK <- colSums(RPK)

# Calculate TPM
TPM <- sweep(RPK, 2, sum_RPK, FUN = "/") * 1e6

# Bind gene id to TPM
TPM_df <- cbind(gene_id = counts_length$gene_id, TPM)

# Log transform 
TPM_df[,-1] <- log2(TPM_df[,-1] + 1)

# Save as csv
#write.csv(TPM_df, "../data/Molecular/mRNA/TPM_log.csv")
TPM_df <- read.csv("../data/Molecular/mRNA/TPM_log.csv")[-1]
```

Look at TPM in 1hpf samples 
```{r}
TPM_df_1hpf <- TPM_df[,1:5]
TPM_df_1hpf_greater1 <- TPM_df_1hpf[rowSums(TPM_df_1hpf[,-1]) > 1, ]
dim(TPM_df_1hpf_greater1)
```

19142 genes were expressed >1 TPM--> the maternal complement. 


Now we want to compare the DEs using the TPMs. We already did the differential expression analysis in [this code](https://github.com/JillAshey/DevelopmentalTimeseries/blob/main/scripts/mRNA_DEG.Rmd). 

Read in metadata for downstream analysis
```{r}
meta <- read.csv("../data/sample_metadata/tube_meta_mcap.csv") %>%
  dplyr::select(TubeID, Date, hpf, Metric, EmbryosPerTube) %>%
  filter(Metric == "Molecular")

# Filter for sequenced samples (and removes the outlier samples)
meta_filt <- meta %>%
  filter(TubeID %in% colnames(counts))
```

## 1 v 4 hpf TPM

Read in 1v4hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v4_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 4hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_4hpf <- de_results_tpm[, c("M10", "M11", "M13")]
TPM_4hpf$avg_4hpf <- rowMeans(TPM_4hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_4hpf = TPM_4hpf$avg_4hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_4hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_4hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (4 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 9 hpf TPM

Read in 1v9hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v9_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 9hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_9hpf <- de_results_tpm[, c("M23", "M24", "M26", "M26")]
TPM_9hpf$avg_9hpf <- rowMeans(TPM_9hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_9hpf = TPM_9hpf$avg_9hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_9hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_9hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (9 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 14 hpf TPM

Read in 1v14hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v14_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 14hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_14hpf <- de_results_tpm[, c("M35", "M36", "M37", "M39")]
TPM_14hpf$avg_14hpf <- rowMeans(TPM_14hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_14hpf = TPM_14hpf$avg_14hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_14hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_14hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (14 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 22 hpf TPM

Read in 1v22hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v22_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 22hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_22hpf <- de_results_tpm[, c("M47", "M48", "M51", "M52")]
TPM_22hpf$avg_22hpf <- rowMeans(TPM_22hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_22hpf = TPM_22hpf$avg_22hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_22hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_22hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (22 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 28 hpf TPM

Read in 1v28hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v28_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 28hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_28hpf <- de_results_tpm[, c("M60", "M62", "M63")]
TPM_28hpf$avg_28hpf <- rowMeans(TPM_28hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_28hpf = TPM_28hpf$avg_28hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_28hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_28hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (28 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 48 hpf TPM

Read in 1v48hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v48_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 48hpf and average across replicates. Plot by TPM
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_48hpf <- de_results_tpm[, c("M72", "M73", "M74", "M75")]
TPM_48hpf$avg_48hpf <- rowMeans(TPM_48hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_48hpf = TPM_48hpf$avg_48hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_48hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_48hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (48 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

## 1 v 72 hpf TPM

Read in 1v72hpf DE results and join with TPM df
```{r}
de_results <- read.csv("../output/Molecular/mRNA/mRNA_DE_results_all_1v72_wald.csv")
colnames(de_results)[1] <- "gene_id"

de_results_tpm <- de_results %>%
  full_join(TPM_df, by = "gene_id")
```

Subset results from 1 and 72hpf and average across replicates
```{r}
TPM_1hpf <- de_results_tpm[, c("gene_id", "M6", "M7", "M8")]
TPM_1hpf$avg_1hpf <- rowMeans(TPM_1hpf[,-1])

TPM_4hpf <- de_results_tpm[, c("M85", "M86", "M87", "M88")]
TPM_4hpf$avg_4hpf <- rowMeans(TPM_4hpf)

plot_df <- data.frame(
  gene_id = TPM_1hpf$gene_id,
  avg_1hpf = TPM_1hpf$avg_1hpf,
  avg_4hpf = TPM_4hpf$avg_4hpf, 
  LFC = de_results_tpm$log2FoldChange,
  padj = de_results_tpm$padj
)

plot_df$DE_status <- ifelse(plot_df$padj < 0.05 &
                           (plot_df$avg_1hpf > 1 | plot_df$avg_4hpf > 1), "DE", "Not DE")

ggplot(plot_df, aes(x = avg_1hpf, y = avg_4hpf, color = DE_status)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_color_manual(values = c("DE" = "orange", "Not DE" = "gray70")) +
  labs(
    x = "Log2 TPM (1 hpf)",
    y = "Log2 TPM (72 hpf)",
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```




