---
title: "WGCNA miRNA"
author: "Jill Ashey"
date: "2025-02-25"
output: html_document
---

WGCNA for Mcap 2023 developmental timeseries lncRNA data. Using code from [Erin Chille](https://github.com/echille/Mcapitata_Developmental_Gene_Expression_Timeseries/blob/master/2a-WGCNA/Developmental_WGCNA.Rmd) and [Zoe Dellaert](https://github.com/imkristenbrown/Heron-Pdam-gene-expression/blob/master/BioInf/scripts/WGCNA/WGCNA.Rmd).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# The following setting is important, do not omit.
options(stringsAsFactors = FALSE) #Set Strings to character
```

Import necessary libraries
```{r, message=FALSE, warning=FALSE}
# install.packages("BiocManager")
# library("BiocManager")
# BiocManager::install("impute", type = "source")
# BiocManager::install("WGCNA",force = TRUE)
# BiocManager::install("vsn")
# install.packages("dendextend")

library(dplyr)
library(tidyr)
library(Rmisc)
library(lubridate)
library(reshape2)
library(ggplot2)
library(car)
library(MASS)
library(mgcv)
#library(MuMIn)
library(emmeans)
#library(locfit)
library(DESeq2)
library(impute)
library(WGCNA)
library(xfun)
library(genefilter)
library(vsn)
library(RColorBrewer)
library(pheatmap)
library(dendextend)
library(ComplexHeatmap)
library(purrr)
sessionInfo()
```

Read in miRNA count data. This data was already normalized by mirdeep2
```{r}
mirna_counts <- read.delim("../output/Molecular/smRNA/mirdeep2/trim_stringent_first_batch/miRNAs_expressed_all_samples_1740330261.csv", header = T)
mirna_counts_df <- as.data.frame(mirna_counts)
mirna_counts_df <- unique(mirna_counts_df)

# Select specific columns 
mirna_counts_df <- mirna_counts_df %>%
  dplyr::select(X.miRNA, s09.norm., s13.norm., s23.norm., s35.norm., s52.norm., s60.norm., s72.norm., s85.norm.)

# Collapse duplicate miRNAs
mirna_counts_df <- mirna_counts_df %>%
  dplyr::group_by(X.miRNA) %>%
  dplyr::summarise(across(ends_with(".norm."), mean))

# Rename cols based on hpf 
colnames(mirna_counts_df) <- c("mirna", "1_hpf", "4_hpf", "9_hpf", "14_hpf", "22_hpf", "28_hpf", "48_hpf", "72_hpf")

# Rename based on sample name
#colnames(mirna_counts_df) <- c("M9", "M13", "M23", "M35", "M52", "M60", "M72", "M85")

# Set row names
mirna_counts_df <- mirna_counts_df %>%
  column_to_rownames("X.miRNA")
```

To run WGCNA, I need more replicates than n=1. To create this, I am going to take each existing column and create 3 new columns (ie M9_1, M9_2, M9_3 etc). For each new column, I will: 
- For "_1" columns: Multiply the original value by a random number between 0.9 and 1.1
- For "_2" columns: Multiply the original value by a random number between 0.8 and 1.2
- For "_3" columns: Multiply the original value by a random number between 0.7 and 1.3

This will introduce some variability without super different values. The new values will be within +- 10%, 20%, and 30% respectively. Initially, I tried to do +1, +2, and +3 to the values but this did not add enough variation to move forward with vst. 
```{r}

set.seed(123)  # for reproducibility

result <- mirna_counts_df %>%
  mutate(across(everything(), 
                list(
                  "1" = ~.x * runif(n(), 0.9, 1.1),
                  "2" = ~.x * runif(n(), 0.8, 1.2),
                  "3" = ~.x * runif(n(), 0.7, 1.3)
                ),
                .names = "{.col}_{.fn}")) %>%
  select(unlist(map(names(mirna_counts_df), ~c(.x, paste0(.x, "_1"), paste0(.x, "_2"), paste0(.x, "_3")))))
```

Convert to integers 
```{r}
numeric_cols <- sapply(result, is.numeric)
result[numeric_cols] <- lapply(result[numeric_cols], function(x) as.integer(round(x, 0)))
str(result)
```

Read in metadata 
```{r}
meta <- read.csv("../data/sample_metadata/tube_meta_mcap.csv") %>%
  dplyr::select(TubeID, Date, hpf, Metric, EmbryosPerTube) %>%
  dplyr::filter(Metric == "Molecular")

meta$hpf <- gsub(" ", "_", meta$hpf)
meta$hpf <- paste0("hpf_", sub("_hpf", "", meta$hpf))

# Filter for sequenced samples
meta_filt <- meta %>%
  filter(TubeID %in% colnames(mirna_counts_df))
```

Create new rows in meta_filt df that match with the simulated data 
```{r}
# Function to create additional rows
create_additional_rows <- function(row) {
  tibble(
    TubeID = paste0(row$TubeID, c("", "_1", "_2", "_3")),
    Date = rep(row$Date, 4),
    hpf = rep(row$hpf, 4),
    Metric = rep(row$Metric, 4),
    EmbryosPerTube = rep(row$EmbryosPerTube, 4),
    original_order = rep(row$original_order, 4)
  )
}

# Add an order column, apply the function, then sort
meta_filt_expanded <- meta_filt %>%
  mutate(original_order = row_number()) %>%
  group_by(TubeID) %>%
  do(create_additional_rows(.)) %>%
  ungroup() %>%
  arrange(original_order, TubeID) %>%
  select(-original_order)
```

Reorder columns in filtered counts matrix so that it matches the metadata
```{r}
result <- result[, meta_filt_expanded$TubeID]

# Sanity check! 
meta_filt_expanded$TubeID %in% colnames(result) #are all of the sample names (rows) in the metadata df in the gene count matrix? Should be TRUE. 
all(rownames(meta_filt_expanded$TubeID) == colnames(result)) #are they the same in the same order? Should be TRUE
```

Set levels of factors 
```{r}
meta_filt_expanded$hpf <- factor(meta_filt_expanded$hpf, levels = c("hpf_1", "hpf_4", "hpf_9", "hpf_14", "hpf_22", "hpf_28", "hpf_48", "hpf_72"))
```

Construct DESeq data set
```{r}
gdds <- DESeqDataSetFromMatrix(countData = result, colData = meta_filt_expanded, design = ~hpf)
```

To do this we first need to calculate the size factors of our samples. This is a rough estimate of how many reads each sample contains compared to the others. In order to use VST (the faster log2 transforming process) to log-transform our data, the size factors need to be less than 4. Otherwise, there could be artefacts in our results.
```{r}
SF.gdds <- estimateSizeFactors(gdds) #estimate size factors to determine if we can use vst  to transform our data. Size factors should be less than for to use vst
print(sizeFactors(SF.gdds)) #View size factors
```

Size factors are all less than 4, so we can use VST!
```{r}
gvst <- varianceStabilizingTransformation(gdds, blind=FALSE) #apply a variance stabilizing transforamtion to minimize effects of small counts and normalize wrt library size
#head(assay(gvst), 3) #view transformed gene count data
```

Plot a heatmap of sample-to-sample distances
```{r}
gsampleDists <- dist(t(assay(gvst))) #calculate distance matix
gsampleDistMatrix <- as.matrix(gsampleDists) #distance matrix
rownames(gsampleDistMatrix) <- colnames(gvst) #assign row names
colnames(gsampleDistMatrix) <- NULL #assign col names
pheatmap(gsampleDistMatrix, #plot matrix
         clustering_distance_rows=gsampleDists, #cluster rows
         clustering_distance_cols=gsampleDists) #cluster columns
```

Compile WGCNA dataset 

Transpose the filtered gene count matrix so that the gene IDs are rows and the sample IDs are columns.
```{r}
datExpr <- as.data.frame(t(assay(gvst))) #transpose to output to a new data frame with the column names as row names. And make all data numeric
```

Look for outliers by examining sample clusters
```{r}
sampleTree = hclust(dist(datExpr), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
# The user should change the dimensions if the window is too large or too small.
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

Network construction and consensus module detection

First choose a soft-threshold power. The soft thresholding power (β) is the number to which the co-expression similarity is raised to calculate adjacency. The function pickSoftThreshold performs a network topology analysis. The user chooses a set of candidate powers, however the default parameters are suitable values.
```{r, message=FALSE, warning=FALSE}
##Soft threshold
dim(datExpr) #  32 x 57
# Choose a set of soft-thresholding powers
powers = c(1:10, seq(by=0.5))
# Call the network topology analysis function

#the below takes a long time to run, so is commented out and the pre-run results are loaded in below. I need to save it to my desktop because it is too large to push to github. Instead, it will go on OSF
sft <- pickSoftThreshold(datExpr, powerVector = powers, verbose = 5) #...wait for this to finish
save(sft, file = "~/Desktop/PutnamLab/DevelopmentalTimeseries/sft_miRNA.RData")
#load("~/Desktop/PutnamLab/DevelopmentalTimeseries/sft_miRNA.RData")

# pickSoftThreshold 
#  performs the analysis of network topology and aids the
# user in choosing a proper soft-thresholding power.
# The user chooses a set of candidate powers (the function provides suitable default values)
# function returns a set of network indices that should be inspected

sizeGrWindow(9, 5) # set window size 
# png to output 
png("../output/Molecular/smRNA/WGCNA/sft.png", 1000, 1000, pointsize=20)
par(mfrow = c(1,2));
cex1 = 0.9;
# # Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
    main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
    labels=powers,cex=cex1,col="red");
# # this line corresponds to using an R^2 cut-off
abline(h=0.7,col="red")
# # Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
    xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
    main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
dev.off() # output 
#I used a scale-free topology fit index **R^2 of 0.85**. This lowest recommended R^2 by Langfelder and Horvath is 0.8. I chose 0.85 because we want to use the smallest soft thresholding power that maximizes with model fit. It appears that our **soft thresholding power is 2** because it is the loweest power before the R^2=0.8 threshold that maximizes with model fit.

plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)", ylab="Scale Free Topology Model Fit,signed R^2",
     main = "Scale independence", ylim = c(-0.5, 0.1))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers, cex=0.9, col="red")
abline(h=max(-sign(sft$fitIndices[,3])*sft$fitIndices[,2]), col="red")

```

Plots look a little wonky since I am simulating the data. My max R^2 is 0.1 lol. 

## Step wise module construction 

#### Step 1: Create adjacency matrix 

```{r}
softPower = 10 # set the soft threshold based on the plots above 

# signed 
adjacency_sign = adjacency(datExpr, power = softPower, type="signed")  #Calculate adjacency
```

###  Step 2: Turn adjacency into topological overlap: Calculation of the topological overlap matrix, (TOM) and the corresponding dissimilarity, from a given adjacency matrix.

```{r}
#the below takes a long time to run, so is commented out and the pre-run results are loaded in below
TOM_sign_mirna = TOMsimilarity(adjacency_sign, TOMType="signed") #Translate adjacency into topological overlap matrix
save(TOM_sign_mirna, file = "~/Desktop/PutnamLab/DevelopmentalTimeseries/TOM_sign_miRNA.Rdata")
#load("~/Desktop/PutnamLab/DevelopmentalTimeseries/TOM_sign_lncRNA.Rdata")

dissTOM_sign   = 1-TOM_sign_mirna
```

###  Step 3: Call the hierarchical clustering function - plot the tree

```{r}
# Call the hierarchical clustering function
#to get this to work, I had to increase my memory limit using: memory.limit(size = 45000) 
geneTree_sign   = hclust(as.dist(dissTOM_sign), method = "average");

# Plot the resulting clustering tree (dendrogram) Each leaf corresponds to a gene, branches grouping together densely are interconnected, highly co-expressed genes.  
sizeGrWindow(12,9)

plot(geneTree_sign, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity - SIGNED",
     labels = FALSE, hang = 0.04)
```


###  Step 4: Set module size and 'cutreeDynamic' to create clusters 

```{r}
#Module identification is essentially cutting the branches off the tree in the dendrogram above. We like large modules, so we set the **minimum module size** relatively high, so we will set the minimum size at 30. For my miRNAs, since there are only 57, I am setting min module size to 2
minModuleSize = 2; # set this for the subseqent call...

dynamicMods_sign = cutreeDynamic(dendro = geneTree_sign, distM = dissTOM_sign,
                            deepSplit = 1, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize);
table(dynamicMods_sign) # number of genes per module. Module 0 is reserved for unassigned genes. The are other modules will be listed largest to smallest. 
```

###  Step 5: convert numeric network to colors and plot the dendrogram

```{r} 
# Convert numeric lables into colors
dynamicColors_sign = labels2colors(dynamicMods_sign) # add colors to module labels (previously numbers)
table(dynamicColors_sign) # lets look at this table...
# Plot the dendrogram and colors underneath

plotDendroAndColors(geneTree_sign, dynamicColors_sign, "Dynamic Tree Cut - SIGNED",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors 'SIGNED'")

sizeGrWindow(8,6)
png("../output/Molecular/smRNA/WGCNA/GeneDendrogram.png", 1000, 1000, pointsize=20)
plotDendroAndColors(geneTree_sign, dynamicColors_sign, "Dynamic Tree Cut - SIGNED",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors 'SIGNED'")
dev.off()
```

###  Step 6: Calculate Eigengenes - view thier connectivity based on 'MEDiss = 1-cor(MEs)'

```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors_sign, softPower = 14)

MEs = MEList$eigengenes
MEs

library(dplyr)
MEs <- MEs %>% 
    select_if(~ !any(is.na(.)))

# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs, use = "pairwise.complete.obs");

# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
png("../output/Molecular/smRNA/WGCNA/ClusterEigengenes.png", 1000, 1000, pointsize=20)
plot(METree, main = "Clustering of module eigengenes - SIGNED (dissimilarity calc = MEDiss = 1-cor(MEs))",
     xlab = "", sub = "")
MEDissThres = 0.15 
abline(h=MEDissThres, col = "red")
dev.off()
```

###  Step 7: Specify the cut line for the dendrogram (module) - Calc MODULE EIGENGENES (mergeMEs)

```{r}
MEDissThres = 0.15 # **Merge modules with >85% eigengene similarity.** Most studies use somewhere between 80-90% similarity. I will use 85% similarity as my merging threshold.
# Plot the cut line into the dendrogram
#abline(h=MEDissThres, col = "red")
# Call an automatic merging function
# merge = mergeCloseModules(dds.d0_vst, dynamicColors, cutHeight = MEDissThres, verbose = 3)
merge = mergeCloseModules(datExpr, dynamicColors_sign, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;
library(dplyr)
mergedMEs <- mergedMEs %>% 
    select_if(~ !any(is.na(.)))
# Cluster module eigengenes
MEDiss2 = 1-cor(mergedMEs,use = 'pairwise.complete.obs');
MEDiss2
METree2 = hclust(as.dist(MEDiss2), method = "average");
# Plot the result
plot(METree2, main = "Clustering of module eigengenes - SIGNED (dissimilarity calc = MEDiss = 1-cor(MEs))",
     xlab = "", sub = "")


sizeGrWindow(7, 6)
png("../output/Molecular/smRNA/WGCNA/ClusterEigengenes_merged.png", 1000, 1000, pointsize=20)
plot(METree2, main = "Clustering of module eigengenes - SIGNED (dissimilarity calc = MEDiss = 1-cor(MEs))",
     xlab = "", sub = "")
dev.off()
```

###  Step 8: Plot dendrogram with the cut line 'MEDissThres' 

```{r}
plotDendroAndColors(geneTree_sign, cbind(dynamicColors_sign, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)

sizeGrWindow(12, 9)

png("../output/Molecular/smRNA/WGCNA/ClusterDendrogram_signed.png", 1000, 1000, pointsize=20)
plotDendroAndColors(geneTree_sign, cbind(dynamicColors_sign, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
dev.off()
```

###  Step 9: Commit to mergedcolors as 'MEs' and 'moduleColors'
```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs;
# Save module colors and labels for use in subsequent parts
save(MEs, moduleLabels, moduleColors, geneTree_sign, file = "~/Desktop/PutnamLab/DevelopmentalTimeseries/networkConstruction-stepByStep_miRNA.RData")
# write csv - save the module eigengenes
write.csv(MEs, file = "../output/Molecular/smRNA/WGCNA/WGCNA_ModuleEigengenes.csv")
table(mergedColors)
```

## Prepare for  module trait associations - Eigengene calc - trait data as factors
```{r}
#Prepare trait data. Data has to be numeric, so I will substitute hpf for numeric values. The "trait" we are considering here is hpf. Make a dataframe that has a column for each hpf and a row for samples. Populate a 1 for samples that match a specific hpf and a 0 for samples not matching. This process changes hpf from a categorical variable into a binary variable. This will allow for correlations between mean eigengenes and habitat of origin.  
allTraits <- names(meta_filt$hpf)
allTraits$hpf_1 <- c(1, 1, 1, 1, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_4 <- c(0, 0, 0, 0, # 1hpf
                    1, 1, 1, 1, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_9 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    1, 1, 1, 1, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_14 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    1, 1, 1, 1, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_22 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    1, 1, 1, 1, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_28 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    1, 1, 1, 1, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_48 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    1, 1, 1, 1,# 48hpf
                    0, 0, 0, 0) # 72hpf
allTraits$hpf_72 <- c(0, 0, 0, 0, # 1hpf
                    0, 0, 0, 0, # 4hpf
                    0, 0, 0, 0, # 9hpf
                    0, 0, 0, 0, # 14hpf
                    0, 0, 0, 0, # 22hpf
                    0, 0, 0, 0, # 28hpf
                    0, 0, 0, 0, # 48hpf
                    1, 1, 1, 1) # 72hpf
datTraits <- as.data.frame(allTraits)
dim(datTraits)
rownames(datTraits) <- meta_filt_expanded$TubeID
print(datTraits)
```

Define numbers of genes and samples
```{r}
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
```

Recalculate MEs with color labels
```{r}
MEs0 = moduleEigengenes(datExpr, moduleColors,softPower=14)$eigengenes
MEs = orderMEs(MEs0)
names(MEs)
```

Correlations of traits with eigengenes
```{r}
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
Colors=sub("ME","",names(MEs))

# Identify genes significantly associated with traits 
geneTraitSignificance = as.data.frame(cor(datExpr, datTraits, use = "p"))
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))

# Combine gene info with module membership and trait correlations 
geneInfo = data.frame(gene = names(datExpr),
                      moduleColor = moduleColors,
                      geneTraitSignificance,
                      GSPvalue)

# Count number of genes in each module 
num_genes <- geneInfo %>%
  dplyr::group_by(moduleColor) %>%
  dplyr::summarise(num_queries = n_distinct(gene))

moduleTraitTree = hclust(dist(t(moduleTraitCor)), method = "average");
plot(moduleTraitTree, main = "Life stage clustering based on module-trait correlation", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

Correlations of genes with eigengenes
```{r}
moduleGeneCor=cor(MEs,datExpr)
moduleGenePvalue = corPvalueStudent(moduleGeneCor, nSamples)
```

Attempt at complexHeatmap
```{r}
#bold sig p-values
#dendrogram with WGCNA MEtree cut-off
#colored y-axis

METree <- hclust(as.dist(1-cor(t(moduleTraitCor))), method = "average")

#Create list of pvalues for eigengene correlation with specific life stages
heatmappval <- signif(moduleTraitPvalue, 1)

#Make list of heatmap row colors
htmap.colors <- names(MEs)
htmap.colors <- gsub("ME", "", htmap.colors)

# Set so that life stages are in order 
hpf_order <- c("hpf_1", "hpf_4", "hpf_9", "hpf_14", "hpf_22", "hpf_28", "hpf_48", "hpf_72")

pdf(file = "../output/Molecular/smRNA/WGCNA/Module-trait-relationship-heatmap.pdf", height = 11.5, width = 8)
ht=Heatmap(moduleTraitCor, name = "Eigengene", column_title = "Module-Trait Eigengene Correlation", 
        col = blueWhiteRed(50), 
        row_names_side = "left", row_dend_side = "left",
        width = unit(4, "in"), height = unit(8.5, "in"), 
        column_order = match(hpf_order, colnames(moduleTraitCor)),
        column_dend_reorder = FALSE, 
        column_split = c(1,1,1,2,2,2,3,3), column_dend_height = unit(0.5, "in"),
        cluster_rows = METree, row_split = 17, row_gap = unit(2.5, "mm"), border = TRUE,
        cell_fun = function(j, i, x, y, w, h, col) {
        if(heatmappval[i, j] <= 0.05) {
            grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 8, fontface = "bold"))
        }
        else {
            grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 8, fontface = "plain"))
        }},
        column_names_gp =  gpar(fontsize = 10),
row_names_gp = gpar(fontsize = 10, alpha = 0.75, border = TRUE, fill = htmap.colors))
draw(ht)
dev.off()
```

```{r}
MEs <- read.csv("../output/Molecular/smRNA/WGCNA/WGCNA_ModuleEigengenes.csv")
MEs_table <- MEs # new table for plotting 
#MEs_table$X <- paste0("M", MEs_table$X)
#MEs_table$TubeID <- rownames(MEs) # call rows as coolumn to merge with treatment data
MEsPlotting <- MEs_table %>%
  full_join(meta_filt_expanded, by = c("X" = "TubeID")) %>%
  dplyr::select(-c("Date", "Metric", "EmbryosPerTube"))
```

Plot by module 
```{r}
MEsPlotting_long <- MEsPlotting %>%
  pivot_longer(cols = -c(X, hpf), names_to = "Module", values_to = "Eigengene")

# Create the faceted boxplot
ggplot(MEsPlotting_long, aes(x = hpf, y = Eigengene, fill = hpf)) +
  geom_boxplot() +
  facet_wrap(~ Module, scales = "free_y", ncol = 4) +  # Adjust ncol as needed
  labs(title = "Module Eigengenes Across Time",
       x = "Hours Post Fertilization",
       y = "Eigengene Value") +
  geom_hline(
    yintercept = 0,
    linetype = 'dotted',
    color = 'black',
    size = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  scale_x_discrete(limits = c("hpf_1", "hpf_4", "hpf_9", "hpf_14", "hpf_22", "hpf_28", "hpf_48", "hpf_72")) +
  scale_fill_brewer(palette = "Set2") +  # You can change the color palette as needed
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgrey", color = NA),
    strip.text = element_text(face = "bold")
  )
ggsave("../output/Molecular/smRNA/WGCNA/module_boxplot.pdf", plot = last_plot(), height = 15, width = 25)
ggsave("../output/Molecular/smRNA/WGCNA/module_boxplot.png", plot = last_plot(), height = 15, width = 25)
```

Use these to determine if maternal, mzt1, mzt2, zga1, or zga2

Maternal = blue
MZT1 = yellow
MZT2 = brown, green
ZGA1 = turquoise

Make ME boxplots by hpf 
```{r}
MEs <- read.csv("../output/Molecular/smRNA/WGCNA/WGCNA_ModuleEigengenes.csv")
MEs_table <- MEs # new table for plotting 
#MEs_table$X <- paste0("M", MEs_table$X)
#MEs_table$TubeID <- rownames(MEs) # call rows as coolumn to merge with treatment data
MEsPlotting <- MEs_table %>%
  full_join(meta_filt_expanded, by = c("X" = "TubeID")) %>%
  dplyr::select(-c("Date", "Metric", "EmbryosPerTube"))

# Average modules to form the following groups: Maternal, MZT1, MZT2, ZGA1, ZGA2
## Group 1: Maternal
maternal <- MEsPlotting %>%
  dplyr::select(MEblue) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Maternal_avg = mean(c_across(everything())))

## Group 2: Wave 1 MZT
wave1_mzt <- MEsPlotting %>%
  dplyr::select(MEyellow) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Wave1_MZT_avg = mean(c_across(everything())))

## Group 3: Wave 2 MZT
wave2_mzt <- MEsPlotting %>%
  dplyr::select(MEbrown, MEgreen) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Wave2_MZT_avg = mean(c_across(everything())))

## Group 4: ZGA1
zga1 <- MEsPlotting %>%
  dplyr::select(MEturquoise) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(ZGA1_avg = mean(c_across(everything())))

## Group 5: ZGA2
# zga2 <- MEsPlotting %>%
#   dplyr::select(MEred, MEcyan, MEsalmon) %>%
#   dplyr::rowwise() %>%
#   dplyr::mutate(ZGA2_avg = mean(c_across(everything())))

# Combine all averages
MEsPlotting_avg <- bind_cols(
  MEsPlotting %>% dplyr::select(X, hpf),
  maternal %>% dplyr::select(Maternal_avg),
  wave1_mzt %>% dplyr::select(Wave1_MZT_avg),
  wave2_mzt %>% dplyr::select(Wave2_MZT_avg),
  zga1 %>% dplyr::select(ZGA1_avg)
  #zga2 %>% dplyr::select(ZGA2_avg)
)

MEsPlotting_melt <- melt(MEsPlotting_avg, id.vars = c("X", "hpf"))
MEsPlotting_melt$hpf <- factor(MEsPlotting_melt$hpf, levels = c("hpf_1", "hpf_4", "hpf_9", "hpf_14", "hpf_22", "hpf_28", "hpf_48", "hpf_72"))

# Plot as boxplots 
MEplot <- ggplot(MEsPlotting_melt, aes(x = hpf, y = value, fill = factor(hpf))) +
  geom_boxplot(
    aes(middle = mean(value)),
    position = position_dodge(0.8),
    outlier.shape = NA,
    alpha = 0.7
  ) + 
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 19,
    size = 3,
    color = "black",
    position = position_dodge(0.75)
  ) +
  geom_hline(
    yintercept = 0,
    linetype = 'dotted',
    color = 'black',
    size = 0.5
  ) +
  scale_y_continuous(
    name = "Module Eigengene",
    limits = c(-0.5, 0.5),
    breaks = seq(-0.5, 0.5, 0.25)
  ) +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~ variable, scales = "free_y") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgrey", color = NA),
    strip.text = element_text(face = "bold")
  ); MEplot
ggsave("../output/Molecular/smRNA/WGCNA/MEs_boxplot.pdf", plot = MEplot, height = 15, width = 25)
ggsave("../output/Molecular/smRNA/WGCNA/MEs_boxplot.png", plot = MEplot, height = 15, width = 25)

# Plot as line plot 
MEplot_line <- ggplot(MEsPlotting_melt, aes(x = factor(hpf), y = value, group = 1)) +
  geom_smooth(
    method = "loess",
    se = TRUE,
    linewidth = 5,  # Increase line thickness
    color = "blue",
    fill = "lightblue",
    alpha = 0.2
  ) +
  # geom_point(
  #   stat = "summary",
  #   fun = "mean",
  #   size = 2,
  #   color = "darkblue"
  # ) +
  # geom_hline(
  #   yintercept = 0,
  #   linetype = 'dotted',
  #   color = 'black',
  #   size = 1
  # ) +
  scale_y_continuous(
    name = "Module Eigengene",
    limits = c(-0.25, 0.25),
    breaks = seq(-0.5, 0.5, 0.25)
  ) +
  scale_x_discrete(
    name = "Hours post fertilization",
    labels = function(x) gsub("hpf_", "", x)  # This removes the "hpf_" prefix
  ) +
  facet_wrap(~ variable, scales = "free_y") +
  theme_bw(base_size = 20) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgrey", color = NA),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(hjust = 1)
  ); MEplot_line

ggsave("../output/Molecular/smRNA/WGCNA/MEs_lineplot.pdf", plot = MEplot_line, height = 15, width = 25)
ggsave("../output/Molecular/smRNA/WGCNA/MEs_lineplot.png", plot = MEplot_line, height = 15, width = 25)
```


